---
title: "Analysis of Ewing surival data for YB-1"
output:
  html_notebook:
      code_folding: none
---

This document describes exploration of some osteosarcoma expression data from the TARGET-OS study in order to look at whether there is any relationship between YB-1 and survival.

## Linked analysis

This file is a continuation of the work done in:

* (20200529_explorationYbx1DepMap)[https://github.com/chrishuges/projectsWorkspace/tree/master/sorensenLab/relatedToYbx1/20200529_explorationYbx1DepMap]

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('DESeq2')
library('jsonlite')
library('org.Hs.eg.db')
library('AnnotationDbi')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

Lastly, we need to call a source file that has some functions we will need. For simplicity and ease of tracking, I make a copy of this file in each directory that I will work in in order to be able to trace the exact one functions that are used, in case I change them and mistakenly don't push it to GitHub.

```{r}
source(paste(baseWorkspace, '/20200601_explorationYbx1SurvivalOsTarget/userDefinedFunctions.R', sep = ''))
```

## Data processing

Althought I don't end up processing the data this way (because I discovered this after), there is some great info (here)[https://www.bioconductor.org/packages/devel/bioc/vignettes/TCGAbiolinks/inst/doc/download_prepare.html] about getting data programmatically from R. Also some good info (here)[https://www.biostars.org/p/204092/].

First I need to get the location of the expression data and get set up to read it.

```{r}
######################################################################################
osDirs = list.dirs(paste(baseRepository, '/20200601_explorationYbx1SurvivalOsTarget', sep = ''), full.names = TRUE)
osDirsSub = osDirs[grepl('explorationYbx1SurvivalOsTarget\\/', osDirs)]
```

I also need to get the annotation data correct to line it up with the clincal data. This took a bit of searching, but I ended up following the directions posted (here)[https://www.biostars.org/p/284708/], specifically by the user Kevin. You need to download a json file of your cases specifically from the Repository section.

```{r}
######################################################################################
manifest = fromJSON(txt = paste(baseRepository, '/20200601_explorationYbx1SurvivalOsTarget/files.2020-06-01.json', sep = ''))
```

Now I can read the data using a function in our userDefinedFunctions file.

```{r, message = FALSE}
######################################################################################
osData = lapply(osDirsSub, countsFileRead, manifest = manifest)
osDataCombine = do.call(rbind, osData)
```

Ok, this looks good. Now we can grab the clinical data from the GDC web portal and read it in here. The columns we are most interested in here are 'case_id', 'days_to_death', and 'vital_status'.

```{r}
######################################################################################
osClinical = read_tsv(paste(baseRepository, '/20200601_explorationYbx1SurvivalOsTarget/clinical.tsv', sep = '')) %>%
  dplyr::select(case_id, days_to_death, vital_status)
colnames(osClinical)[1] = 'sampleName'
osClinical$days_to_death = ifelse(grepl('[0-9]', osClinical$days_to_death), as.numeric(osClinical$days_to_death), NA)
```

Now we want to make it into a DESeq object. If you want, you can get all these files in the same directory manually by going to the directory, typing '*.*' in the search bar, and copying all the htseq files over. I am sure there is a better way to do this, but it isn't necessary for the code below.

```{r}
######################################################################################
directory = paste(baseRepository, '/20200601_explorationYbx1SurvivalOsTarget', sep = '')
sampleTable = data.frame(tibble(sampleName = unique(osDataCombine$caseID),
                         fileName = unique(osDataCombine$dePath)) %>%
  left_join(osClinical) %>%
  mutate(condition = vital_status) %>%
  dplyr::select(sampleName, fileName, condition))
##make the DESeq object
osDE = DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                   directory = directory,
                                   design= ~ condition)
```

Process the data for differential expression using DESeq2 as described (here)[http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html].

```{r}
######################################################################################
keep <- rowSums(counts(osDE)) >= 10
dds <- osDE[keep,]
ddsRes <- DESeq(dds)
res <- results(ddsRes)
resultsNames(ddsRes)
res <- results(ddsRes, name="condition_Dead_vs_Alive")
resOrdered <- res[order(res$pvalue),]
summary(res)
write.csv(as.data.frame(resOrdered), 
          file = paste(baseRepository, '/20200601_explorationYbx1SurvivalOsTarget/deseqOutput.csv', sep = ''))
```




















Now, we will only keep YBX1, which is ENSG00000065978. You could map all the ensg IDs to symbol using org.Hs database, but I only need 1 gene, so this is just more simple.

```{r}
######################################################################################
geneData = osDataAnnotated %>% 
  filter(ensg == 'ENSG00000065978') %>%
  mutate(symbol = AnnotationDbi::select(org.Hs.eg.db, keys = 'ENSG00000065978', columns = c('ENSEMBL','SYMBOL'), keytype = 'ENSEMBL')$SYMBOL) %>%
  dplyr::select(caseID, ensg, symbol, days_to_death, vital_status, counts)


```





defined as low (25% of samples with lowest expression), high (25% with highest expression), and medium (50% of samples with medium expression)
## Session info

```{r}
sessionInfo()
```


