---
title: "Analysis of PuMa data for MG63.3 GUC clones"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of the output PuMa data from the parental and GUC6 clones.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read the quantified puma data sent from Mike.

```{r}
##########################################################################################
puma = read_tsv(paste(baseRepository, '/puma20201030_mg633Ybx1Guc6Puma/mg633Ybx1Guc6PumaTumorBurdenData.txt', sep = '')) %>%
  mutate(slice = factor(seq(1,8,1), levels = seq(1,8,1))) %>%
  pivot_longer(cols = `Wildtype Day 0`:`GUC6 Day 14`, names_to = 'sampleName', values_to = 'tumorBurden') %>%
  mutate(day = factor(sub('.* Day (.*)$', '\\1', sampleName), levels = c(0,14))) %>%
  mutate(cellLine = factor(sub('(.*) Day .*', '\\1', sampleName), levels = c('Wildtype','GUC6')))
```

<div style="margin-bottom:50px;"></div>

Now we can make a plot of these data.

```{r}
##########################################################################################
ggplot(puma, aes(day, tumorBurden, color = cellLine)) +
  geom_point(size = 2, position = position_jitterdodge(jitter.width = 0.50)) +
  labs(x = 'Time (days)', y = 'Percent lung tumor burden', title = 'YBX1 GUC clone PuMa') +
  scale_y_continuous(limits = c(0, 15), breaks = seq(0,15,3)) +
  scale_color_manual(values = c(brewer.pal(3,'Greys')[2], rep(brewer.pal(3,'OrRd')[3],2))) +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/puma20201030_mg633Ybx1Guc6Puma/scatter_mg633Ybx1GucClonePuma.pdf', sep = ''),
       height = 2,
       width = 2,
       useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Get the p-value for the parental vs guc6 at day 14.

```{r}
##########################################################################################
t.test(tumorBurden ~ cellLine, data = puma[grepl('14',puma$day),])

```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```