---
title: "CHLA cell line proteomics"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of proteomics data acquired for comparison of the CHLA9 and CHLA10 cell lines. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('RColorBrewer')
library('ggplot2')
library('DEqMS')
library('ggrepel')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToOthers'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToOthers'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

This is just a single batch, so we only have a single file we need to read in and process.

```{r, message = FALSE}
##########################################################################################
psmId = read_tsv(paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/psm.tsv', sep = '')) %>%
  dplyr::select(Peptide, `Is Unique`, `Is Used`, `Protein ID`, Gene, chla9_rep1:chla10_rep3, hek293_rep1:hek293_rep3) %>%
  dplyr::rename(sequence = Peptide, 
                unique = `Is Unique`,
                used = `Is Used`,
                accession = `Protein ID`,
                gene = Gene) %>%
  dplyr::filter(unique == TRUE, !grepl('CONT', accession), !is.na(gene), !grepl('-', gene)) %>%
  dplyr::rowwise(sequence, accession, gene) %>%
  dplyr::mutate(totalSignal = sum(c_across(chla9_rep1:hek293_rep3), na.rm = TRUE)) %>%
  dplyr::filter(totalSignal >= 100)
```

<div style="margin-bottom:50px;"></div>

Now we can process these data with DEqMS. 

Our sample layout is:

* 126C - chla9_rep1
* 127N - chla10_rep1
* 127C - chla9_rep2
* 128N - chla10_rep2
* 128C - chla9_rep3
* 129N - chla10_rep3
* 129C - empty
* 130N - empty
* 130C - hek293_rep1
* 131N - hek293_rep2
* 131C - hek293_rep3

We will also normalize the data to deal with any gross loading differences. We save both the non-normalized data here as well and check the normalization with a plot.

```{r}
##########################################################################################
quantInput = psmId[,c(1,5:14)]
saveRDS(quantInput, paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/dataset_peptideQuantDataPreDEqMS.rds', sep = ''))
quantInputLog = quantInput
quantInputLog[,3:11][quantInputLog[,3:11] == 0] = NA
quantInputLog[,3:11] = log2(quantInputLog[,3:11])
quantInputNormalized = medianSweeping(quantInputLog, group_col = 2)
boxplot(quantInputNormalized, las = 2, ylab = 'log2 ratio', main = 'normalized TMT data')
```

<div style="margin-bottom:50px;"></div>

All of the data look tight. Good. Lets continue with the DEqMS analysis. We need to build our sample table.

```{r}
##########################################################################################
cond = as.factor(c('chla9','chla10','chla9','chla10','chla9','chla10','hek293','hek293','hek293'))
design = model.matrix(~0+cond) 
colnames(design) = gsub('cond', '', colnames(design))
```

<div style="margin-bottom:50px;"></div>

Now we can build the Limma model.

```{r}
##########################################################################################
geneMatrix = as.matrix(quantInputNormalized)
limmaFit1 = lmFit(geneMatrix, design)
limmaContrasts = c('chla9 - chla10', 'chla9 - hek293', 'chla10 - hek293')
limmaContrastDesign =  makeContrasts(contrasts = limmaContrasts, levels = design)
limmaFit2 = eBayes(contrasts.fit(limmaFit1, contrasts = limmaContrastDesign))
```

<div style="margin-bottom:50px;"></div>

Now for the actual DEqMS work. We will first

```{r}
##########################################################################################
psmCountTable = as.data.frame(table(quantInput$gene))
rownames(psmCountTable) = psmCountTable$Var1
limmaFit2$count = psmCountTable[rownames(limmaFit2$coefficients),2]
limmaFit3 = spectraCounteBayes(limmaFit2)
head(limmaFit3$coefficients)
##
deqmsResults = outputResult(limmaFit3, coef_col = 1)
dim(deqmsResults)
write.table(deqmsResults, 
            paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/dataset_deqms_chla9-chla10.csv', sep = ''),
            sep = ',', row.names = FALSE, quote = FALSE)
##
deqmsResults = outputResult(limmaFit3, coef_col = 2)
dim(deqmsResults)
write.table(deqmsResults, 
            paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/dataset_deqms_chla9-hek293.csv', sep = ''),
            sep = ',', row.names = FALSE, quote = FALSE)
##
deqmsResults = outputResult(limmaFit3, coef_col = 3)
dim(deqmsResults)
write.table(deqmsResults, 
            paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/dataset_deqms_chla10-hek293.csv', sep = ''),
            sep = ',', row.names = FALSE, quote = FALSE)
```

<div style="margin-bottom:50px;"></div>

Generate the variance plot to see the DEqMS model for the data.

```{r}
##########################################################################################
VarianceBoxplot(limmaFit3, n=20, xlab = 'PSM count', main = 'DEqMS analysis of TMT data')
```

<div style="margin-bottom:50px;"></div>

This looks as it should. Now let us make a volcano plot to show the data.

```{r}
##########################################################################################
deqmsResults = outputResult(limmaFit3, coef_col = 1)
deqmsResults$plotPValue = -log10(deqmsResults$sca.P.Value)
deqmsResults$pColors = ifelse(deqmsResults$logFC <= -0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[8],
                              ifelse(deqmsResults$logFC >= 0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[1],
                                     brewer.pal(8,'Greys')[6]))
deqmsResults$pText = ifelse(grepl('YBX1$',deqmsResults$gene), deqmsResults$gene, '')
##
ggplot(deqmsResults, aes(logFC, plotPValue)) +
  geom_point(size = 1, color = deqmsResults$pColors, alpha = 0.5) +
  geom_text_repel(label = deqmsResults$pText, nudge_x = -2) +
  labs(x = 'log2(chla9 vs chla10)', y = '-log10(P-value)', title = 'CHLA9 vs CHLA10 protein abundance') +
  scale_y_continuous(limits = c(0,12), breaks = seq(0,18,3)) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-6,6,2)) +
  geom_vline(xintercept = c(-0.5,0.5), linetype = 'dashed') +
  geom_hline(yintercept = 3, linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/scatter_deqms_chla9-chla10.pdf', sep = ''),
       height = 4, width = 4, useDingbats = FALSE)


##########################################################################################
deqmsResults = outputResult(limmaFit3, coef_col = 2)
deqmsResults$plotPValue = -log10(deqmsResults$sca.P.Value)
deqmsResults$pColors = ifelse(deqmsResults$logFC <= -0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[8],
                              ifelse(deqmsResults$logFC >= 0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[1],
                                     brewer.pal(8,'Greys')[6]))
deqmsResults$pText = ifelse(grepl('YBX1',deqmsResults$gene), deqmsResults$gene, '')
##
ggplot(deqmsResults, aes(logFC, plotPValue)) +
  geom_point(size = 1, color = deqmsResults$pColors, alpha = 0.5) +
  geom_text_repel(label = deqmsResults$pText, nudge_x = -2) +
  labs(x = 'log2(chla9 vs hek293)', y = '-log10(P-value)', title = 'CHLA9 vs HEK293 protein abundance') +
  scale_y_continuous(limits = c(0,12), breaks = seq(0,18,3)) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-6,6,2)) +
  geom_vline(xintercept = c(-0.5,0.5), linetype = 'dashed') +
  geom_hline(yintercept = 3, linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/scatter_deqms_chla9-hek293.pdf', sep = ''),
       height = 4, width = 4, useDingbats = FALSE)


##########################################################################################
deqmsResults = outputResult(limmaFit3, coef_col = 3)
deqmsResults$plotPValue = -log10(deqmsResults$sca.P.Value)
deqmsResults$pColors = ifelse(deqmsResults$logFC <= -0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[8],
                              ifelse(deqmsResults$logFC >= 0.5 & deqmsResults$plotPValue >= 3, brewer.pal(8,'RdBu')[1],
                                     brewer.pal(8,'Greys')[6]))
deqmsResults$pText = ifelse(grepl('YBX1$',deqmsResults$gene), deqmsResults$gene, '')
##
ggplot(deqmsResults, aes(logFC, plotPValue)) +
  geom_point(size = 1, color = deqmsResults$pColors, alpha = 0.5) +
  geom_text_repel(label = deqmsResults$pText, nudge_x = -2) +
  labs(x = 'log2(chla10 vs hek293)', y = '-log10(P-value)', title = 'CHLA10 vs HEK293 protein abundance') +
  scale_y_continuous(limits = c(0,12), breaks = seq(0,18,3)) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-6,6,2)) +
  geom_vline(xintercept = c(-0.5,0.5), linetype = 'dashed') +
  geom_hline(yintercept = 3, linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20210218_chlaCellLineProteomics/scatter_deqms_chla10-hek293.pdf', sep = ''),
       height = 4, width = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```

