---
title: "Polysome RBPs"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of proteomics data acquired for comparison of RBPs in polysomes. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('RColorBrewer')
library('ggplot2')
library('ggrepel')
library('matrixStats')
library('DEqMS')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read the peptide data. This is a SILAC experiment with the light channel as an untreated, and the heavy as an arsenite treated sample. 

```{r, message = FALSE}
##########################################################################################
samples = tibble('sampleName' = paste('polyRbp_',seq(1,3,1),sep=''),
                 'samplePrefix' = 'polyRbp')

##
peptideFiles = file.path(baseRepository, 'proteomics20210324_hek293PolysomeRbpsSearchOutput', samples$sampleName, 'ion_label_quant.tsv')
all(file.exists(peptideFiles))


##
peptideSet = data.frame()
for (i in 1:length(peptideFiles)){
  tempPeptides = read_tsv(peptideFiles[i]) %>%
    dplyr::select(`Protein ID`, `Gene`, `Peptide Sequence`, `Light Intensity`, `Heavy Intensity`) %>%
    dplyr::rename(accession = `Protein ID`, symbol = `Gene`, sequence = `Peptide Sequence`, 
                  noCL = `Light Intensity`, 
                  CL = `Heavy Intensity`) %>%
    dplyr::mutate(dataset = samples[i,'sampleName']$sampleName) %>%
    dplyr::filter(noCL > 0 | CL > 0)
  ##
  peptideSet = rbind(peptideSet, tempPeptides)
}
saveRDS(peptideSet, 
        paste(baseRepository, '/proteomics20210324_hek293PolysomeRbpsSearchOutput/dataset_allPeptideData.rds', sep = ''))

```

<div style="margin-bottom:50px;"></div>

Now work the peptide set through deqms.

```{r, message = FALSE}
##########################################################################################
pep = readRDS(paste(baseRepository, '/proteomics20210324_hek293PolysomeRbpsSearchOutput/dataset_allPeptideData.rds', sep = '')) %>%
  dplyr::select(sequence, symbol, noCL, CL, dataset) %>%
  dplyr::group_by(sequence, symbol, dataset) %>%
  dplyr::summarise(meanNoCL = mean(noCL, na.rm = TRUE), meanCL = mean(CL, na.rm = TRUE)) %>%
  pivot_longer(cols = meanNoCL:meanCL, names_to = 'treatment', values_to = 'intensity') %>%
  pivot_wider(names_from = c(dataset, treatment), values_from = intensity) %>%
  dplyr::select(sequence, symbol, c(3,5,7,4,6,8)) %>%
  dplyr::filter(!is.na(symbol))

##
pepSub = subset(pep, rowSums(is.na(pep[,3:8])) < 3)
pepSub[,3:8] = log2(pepSub[,3:8])
pepSubNorm = medianSweeping(pepSub, group_col = 2)
boxplot(pepSubNorm, las=2, ylab='log2 ratio', main='normalized data')

##
geneMatrix = as.matrix(pepSubNorm)
cond = as.factor(rep(c('noCL','CL'), each = 3))
design = model.matrix(~0+cond) 
colnames(design) = gsub('cond', '', colnames(design))
fit1 = lmFit(geneMatrix, design)
forContrast = c('noCL-CL')
contrast =  makeContrasts(contrasts = forContrast, levels = design)
fit2 = eBayes(contrasts.fit(fit1, contrasts = contrast))
psmCountTable = as.data.frame(table(pepSub$symbol))
rownames(psmCountTable) = psmCountTable$Var1

fit2$count = psmCountTable[rownames(fit2$coefficients),2]
fit3 = spectraCounteBayes(fit2)
deqmsResults = outputResult(fit3,coef_col = 1) 
head(deqmsResults)
```

<div style="margin-bottom:50px;"></div>

Plot some of the results.

```{r, message = FALSE}
##########################################################################################
plotData = as.data.frame(deqmsResults)
write.table(plotData, paste(baseRepository, '/proteomics20210324_hek293PolysomeRbpsSearchOutput/dataset_deqmsAnalysis_noCL-CL.tsv', sep = ''),
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
plotData$logPValue = -log10(plotData$P.Value)
plotData$logPValueScaled = ifelse(plotData$logPValue > 10, 10, plotData$logPValue)
plotData$logFCScaled = ifelse(plotData$logFC > 6, 6, plotData$logFC)
plotData$logFCScaled = ifelse(plotData$logFC < -6, -6, plotData$logFC)
plotData$pColors = ifelse(plotData$P.Value <= 0.001 & plotData$logFC >= 1, brewer.pal(8,'Spectral')[1], 
                        ifelse(plotData$P.Value <= 0.001 & plotData$logFC <= -1, brewer.pal(8,'Spectral')[8], brewer.pal(11,'RdGy')[8]))
goi = c('YBX1','RPL26','IGF2BP1','ELAVL1','DHX9','SYNCRIP','RPS6','STAU1','HNRNPU')
plotData$pText = ifelse(plotData$gene %in% goi, plotData$gene, '')
##
ggplot(plotData, aes(logFCScaled, logPValueScaled)) +
  geom_point(size = 1, color = plotData$pColors, alpha = 0.75) +
  labs(x = 'log2(noCL - CL)', y = '-log10(P-value)', title = 'Polysome RBPs') +
  geom_text_repel(label = plotData$pText, nudge_x = 2, nudge_y = 2) +
  scale_x_continuous(limits = c(-6,3), breaks = seq(-10,10,2)) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,2)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
##
ggsave(paste(baseRepository, '/proteomics20210324_hek293PolysomeRbpsSearchOutput/scatter_deqmsAnalysis_noCL-CL.pdf', sep = ''),
      height = 4, width = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Process the membrane polysome data just as above.

```{r, message = FALSE}
##########################################################################################
samplePrefix = c('memEdta','memTime0','memTime30','memTime60','memTime120','memTime240')
samples = tibble('sampleName' = paste(rep(samplePrefix,each=3),'_',seq(1,3,1),sep=''),
                 'samplePrefix' = rep(samplePrefix, each=3),
                 'treatment' = c(rep('edta',3),rep('dmso',3),rep('arsenite',12)))

##
peptideFiles = file.path(baseRepository, 'proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3', samples$sampleName, 'peptide_label_quant.tsv')
all(file.exists(peptideFiles))


##
peptideSet = data.frame()
for (i in 1:length(peptideFiles)){
  tempPeptides = read_tsv(peptideFiles[i]) %>%
    dplyr::select(`Protein ID`, `Gene`, `Peptide Sequence`, `Light Intensity`, `Heavy Intensity`) %>%
    dplyr::rename(accession = `Protein ID`, symbol = `Gene`, sequence = `Peptide Sequence`, 
                  light = `Light Intensity`, 
                  heavy = `Heavy Intensity`) %>%
    dplyr::mutate(dataset = samples[i,'sampleName']$sampleName) %>%
    dplyr::filter(light > 0 | heavy > 0)
  ##
  peptideSet = rbind(peptideSet, tempPeptides)
}
saveRDS(peptideSet, 
        paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/dataset_allPeptideDataMembrane.rds', sep = ''))

```

<div style="margin-bottom:50px;"></div>

Now work the peptide set through deqms.

```{r, message = FALSE}
##########################################################################################
pep = readRDS(paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/dataset_allPeptideDataMembrane.rds', sep = '')) %>%
  dplyr::select(sequence, symbol, light, dataset) %>%
  dplyr::group_by(sequence, symbol, dataset) %>%
  dplyr::summarise(mean = mean(light, na.rm = TRUE)) %>%
  pivot_wider(names_from = dataset, values_from = mean) %>%
  dplyr::select(sequence, symbol, samples$sampleName[c(1:3,4:6)]) %>%
  dplyr::filter(!is.na(symbol))

##
pepSub = subset(pep, rowSums(is.na(pep[,3:8])) < 3)
pepSub[,3:8] = log2(pepSub[,3:8])
pepSubNorm = medianSweeping(pepSub, group_col = 2)
boxplot(pepSubNorm, las=2, ylab='log2 ratio', main='normalized data')

##
geneMatrix = as.matrix(pepSubNorm)
cond = as.factor(rep(c('edta','time0'), each = 3))
design = model.matrix(~0+cond) 
colnames(design) = gsub('cond', '', colnames(design))
fit1 = lmFit(geneMatrix, design)
forContrast = c('edta-time0')
contrast =  makeContrasts(contrasts = forContrast, levels = design)
fit2 = eBayes(contrasts.fit(fit1, contrasts = contrast))
psmCountTable = as.data.frame(table(pepSub$symbol))
rownames(psmCountTable) = psmCountTable$Var1

fit2$count = psmCountTable[rownames(fit2$coefficients),2]
fit3 = spectraCounteBayes(fit2)
deqmsResults = outputResult(fit3,coef_col = 1) 
head(deqmsResults)
```

<div style="margin-bottom:50px;"></div>

Plot some of the results.

```{r, message = FALSE}
##########################################################################################
plotData = as.data.frame(deqmsResults)
write.table(plotData, paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/dataset_deqmsAnalysisMembrane_edta-time0.tsv', sep = ''),
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
plotData$logPValue = -log10(plotData$P.Value)
plotData$logPValueScaled = ifelse(plotData$logPValue > 10, 10, plotData$logPValue)
plotData$logFCScaled = ifelse(plotData$logFC > 6, 6, plotData$logFC)
plotData$logFCScaled = ifelse(plotData$logFC < -6, -6, plotData$logFC)
plotData$pColors = ifelse(plotData$P.Value <= 0.001 & plotData$logFC >= 1, brewer.pal(8,'Spectral')[1], 
                        ifelse(plotData$P.Value <= 0.001 & plotData$logFC <= -1, brewer.pal(8,'Spectral')[8], brewer.pal(11,'RdGy')[8]))
plotData$pText = ifelse(grepl('YBX1',plotData$gene), plotData$gene, '')
##
ggplot(plotData, aes(logFCScaled, logPValueScaled)) +
  geom_point(size = 1, color = plotData$pColors, alpha = 0.75) +
  labs(x = 'log2(edta - time0)', y = '-log10(P-value)', title = 'Polysome background') +
  geom_text_repel(label = plotData$pText, nudge_x = 2, nudge_y = -1) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-10,10,2)) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,2)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
##
ggsave(paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/scatter_deqmsAnalysisMembrane_edta-time0.pdf', sep = ''),
      height = 4, width = 4, useDingbats = FALSE)
```





















<div style="margin-bottom:50px;"></div>

Now do it for one of the arsenite samples.

```{r, message = FALSE}
##########################################################################################
pep = readRDS(paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/dataset_allPeptideData.rds', sep = '')) %>%
  dplyr::select(sequence, symbol, light, heavy, dataset) %>%
  dplyr::group_by(sequence, symbol, dataset) %>%
  dplyr::summarise(mean_light = mean(light, na.rm = TRUE), mean_heavy = mean(heavy, na.rm = TRUE)) %>%
  dplyr::filter(grepl('cytoTime120',dataset)) %>%
  pivot_wider(names_from = dataset, values_from = mean_light:mean_heavy) %>%
  dplyr::filter(!is.na(symbol))

##
pepSub = subset(pep, rowSums(is.na(pep[,3:8])) < 3)
pepSub[,3:8] = log2(pepSub[,3:8])
pepSubNorm = medianSweeping(pepSub, group_col = 2)
boxplot(pepSubNorm, las=2, ylab='log2 ratio', main='normalized data')

##
geneMatrix = as.matrix(pepSubNorm)
cond = as.factor(rep(c('time120dmso','time120arsenite'), each = 3))
design = model.matrix(~0+cond) 
colnames(design) = gsub('cond', '', colnames(design))
fit1 = lmFit(geneMatrix, design)
forContrast = c('time120dmso-time120arsenite')
contrast =  makeContrasts(contrasts = forContrast, levels = design)
fit2 = eBayes(contrasts.fit(fit1, contrasts = contrast))
psmCountTable = as.data.frame(table(pepSub$symbol))
rownames(psmCountTable) = psmCountTable$Var1

fit2$count = psmCountTable[rownames(fit2$coefficients),2]
fit3 = spectraCounteBayes(fit2)
deqmsResults = outputResult(fit3,coef_col = 1) 
head(deqmsResults)

##
plotData = as.data.frame(deqmsResults)
write.table(plotData, paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/dataset_deqmsAnalysis_time120dmso-arsenite.tsv', sep = ''),
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
plotData$logPValue = -log10(plotData$P.Value)
plotData$logPValueScaled = ifelse(plotData$logPValue > 10, 10, plotData$logPValue)
plotData$logFCScaled = ifelse(plotData$logFC > 6, 6, plotData$logFC)
plotData$logFCScaled = ifelse(plotData$logFC < -6, -6, plotData$logFC)
plotData$pColors = ifelse(plotData$P.Value <= 0.001 & plotData$logFC >= 1, brewer.pal(8,'Spectral')[1], 
                        ifelse(plotData$P.Value <= 0.001 & plotData$logFC <= -1, brewer.pal(8,'Spectral')[8], brewer.pal(11,'RdGy')[8]))
##
ggplot(plotData, aes(logFCScaled, logPValueScaled)) +
  geom_point(size = 1, color = plotData$pColors, alpha = 0.75) +
  labs(x = 'log2(dmso - control)', y = '-log10(P-value)', title = 'Arsenite polysomes') +
  #geom_text_repel(label = ipData$pText, nudge_x = 2, nudge_y = -1) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-10,10,2)) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,2)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
##
ggsave(paste(baseRepository, '/proteomics20210309_hek293ArseniteTimecourseSearchOutput/fraction3/scatter_deqmsAnalysis_time120dmso-arsenite.pdf', sep = ''),
      height = 4, width = 4, useDingbats = FALSE)

```















<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```

