---
title: "Analysis of survival data for Ewing sarcoma"
output:
  html_notebook:
      code_folding: none
---

I again want to look at survival data for Ewing sarcoma, but instead of just going set by set, I am going to combine all the sets I can find into one object and go from there. The details for each set will be presented as we go through.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('GEOquery')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

We need to call a source file that has some functions we will need. For simplicity and ease of tracking, I make a copy of this file in each directory that I will work in in order to be able to trace the exact one functions that are used, in case I change them and mistakenly don't push it to GitHub.

```{r}
source(paste(baseWorkspace, '/microarray20200604_ewingSarcomaAllSets/userDefinedFunctions.R', sep = ''))
```

## Data processing

This data set is from Postel-Vinay et al. from GSE34620 that was published in PMID: 22327514. I had to contact the authors to get the survival data for this study. First we will get the data.

```{r}
##########################################################################################
esArrayData = getGEO('GSE34620',
             destdir = paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/', sep = ''),
             GSEMatrix = TRUE)
esData = esArrayData[[1]]
```

We now have an expression set and we can look at some of the features to see that we have all the information we need.

```{r}
##################################################################################
head(pData(esData))
head(exprs(esData))[,1:5]
head(fData(esData))
```

First I just want to annotate the expression data a bit to include gene symbols, as these are what I will use for the survival analysis downstream. We can use the feature data to get this annotation.

```{r}
##################################################################################
esExpression = as_tibble(exprs(esData)) %>%
  mutate(ID = row.names(exprs(esData)))
esExpressionAnnotated = as_tibble(fData(esData)) %>%
  mutate(symbol = sapply(strsplit(`Gene Symbol`, ' '), '[', 1)) %>%
  dplyr::select(ID, GB_ACC, symbol) %>%
  left_join(esExpression)
```

Now I want to look at the phenotype data. Some of the information we need is here.

```{r}
##################################################################################
esPhenotype = as_tibble(pData(esData)) %>%
  dplyr::select(geo_accession, characteristics_ch1:characteristics_ch1.2) %>%
  mutate(type = sub('.*type\\: (.*) Tumor$', '\\1', characteristics_ch1.2)) %>%
  mutate(age = as.numeric(sub('age\\: (.*)$', '\\1', characteristics_ch1.1)) * 12) %>%
  mutate(sex = sub('gender\\: (.*)$', '\\1', characteristics_ch1)) %>%
  dplyr::select(geo_accession, type:sex)
```

I need to add the clinical survival data, and these were provided to me directly from the authors.

```{r}
##################################################################################
esPheno = read_tsv(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/Clinical_data_GSE34620.txt', sep = '')) %>%
  mutate(geo_accession = `GEO portal`) %>%
  mutate(ovs = `OS (days)` / 30) %>%
  mutate(efs = `FPS (days)` / 30) %>%
  mutate(status = `statut (1= Death; 0=Life)`) %>%
  dplyr::select(geo_accession:status) %>%
  left_join(esPhenotype) %>%
  dplyr::select(geo_accession, type, age, sex, ovs, efs, status)
esPheno$status = esPheno$status + 1
```

Now we can save the two datasets.

```{r}
##################################################################################
saveRDS(esExpressionAnnotated, paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_postelVinayExpression.rds', sep = ''))
saveRDS(esPheno, paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_postelVinayPhenotype.rds', sep = ''))
```

## Session info

```{r}
sessionInfo()
```