---
title: "Analysis of survival data for Ewing sarcoma"
output:
  html_notebook:
      code_folding: none
---

I again want to look at survival data for Ewing sarcoma, but instead of just going set by set, I am going to combine all the sets I can find into one object and go from there. I have performed the preprocessing of the data, and now will combine and do survival analysis.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('survival')
library('survminer')
library('ggrepel')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

We need to call a source file that has some functions we will need. For simplicity and ease of tracking, I make a copy of this file in each directory that I will work in in order to be able to trace the exact one functions that are used, in case I change them and mistakenly don't push it to GitHub.

```{r}
source(paste(baseWorkspace, '/microarray20200604_ewingSarcomaAllSets/userDefinedFunctions.R', sep = ''))
```

<div style="margin-bottom:50px;"></div>

## Data processing

First we will read in all the data sets.

```{r}
##########################################################################################
savolaExpression = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_savolaExpression.rds', sep = ''))
savolaPhenotype = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_savolaPhenotype.rds', sep = ''))
volchenExpressionA = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_volchenboumExpressionSetA.rds', sep = ''))
volchenPhenotypeA = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_volchenboumPhenotypeSetA.rds', sep = ''))
volchenExpressionB = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_volchenboumExpressionSetB.rds', sep = ''))
volchenPhenotypeB = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_volchenboumPhenotypeSetB.rds', sep = ''))
postelExpression = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_postelVinayExpression.rds', sep = ''))
postelPhenotype = readRDS(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/dataset_postelVinayPhenotype.rds', sep = ''))
```

<div style="margin-bottom:50px;"></div>

Read in the crispr and rna expression data we created previously.

```{r}
##########################################################################################
esHits = readRDS(paste(baseRepository, '/database20200604_depmapEwing/dataset_ewingRnaExprsCrisprDep.rds', sep = ''))
#esHitsGenes = esHits[esHits$medDep <= -1 & esHits$medExprs >= 4, 'symbol']
esHitsGenes = esHits[1:250,]
```

<div style="margin-bottom:50px;"></div>

Do a test survival analysis for genes against the savola data.

```{r, message = FALSE}
##########################################################################################
expressionData = savolaExpression
phenotypeData = savolaPhenotype
geneOfInterest = as.list(esHitsGenes$symbol)
coxZScores = lapply(geneOfInterest, survivalAnalysis, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = FALSE,
                    survivalData = FALSE,
                    printData = FALSE)
esHitsGenes$coxz = unlist(coxZScores)
```

<div style="margin-bottom:50px;"></div>

Now we can make a plot for these scores.

```{r}
##########################################################################################
esHitsPlot = esHitsGenes %>%
  arrange(coxz) %>%
  mutate(index = seq(1,nrow(esHitsGenes),1))
esHitsPlot$pText = ifelse(esHitsPlot$symbol == 'YBX1', esHitsPlot$symbol, '')
esHitsPlot$pSize = ifelse(esHitsPlot$symbol == 'YBX1', 3, 1)
esHitsPlot$pColors = ifelse(esHitsPlot$symbol == 'YBX1', brewer.pal(8,'Spectral')[1], brewer.pal(8,'Greys')[5])
#esHitsPlot$pText = ifelse(esHitsPlot$coxz >= 2.5, esHitsPlot$symbol, '')
##
ggplot(esHitsPlot, aes(coxz, index)) +
  geom_point(size = esHitsPlot$pSize, color = esHitsPlot$pColors) +
  geom_text_repel(label = esHitsPlot$pText, size = 3, segment.size = 0.25) +
  labs(x = 'Cox Univariate Z Score', y = 'Gene index', title = 'Overall survival') +
  theme_classic()
ggsave(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/scatter_savolaOverallSurvivalTop250.pdf', sep = ''), 
       width = 3, height = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Now we can make a nice plot for YB-1 survival.

```{r}
##########################################################################################
expressionData = savolaExpression
phenotypeData = savolaPhenotype
geneOfInterest = 'YBX1'
coxZScores = lapply(geneOfInterest, survivalAnalysis, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = TRUE,
                    survivalData = TRUE,
                    writeDirectory = '/microarray20200604_ewingSarcomaAllSets',
                    printData = TRUE)
```

<div style="margin-bottom:50px;"></div>

Ignore other data sets for now until we see if we can get that really large cohort from Delattre. 

Ok, now we have the data from Delattre, so we can try it out! This is version 2 of the script with a modeled cut off.

```{r, message = FALSE}
##########################################################################################
expressionData = postelExpression
phenotypeData = postelPhenotype
geneOfInterest = as.list(esHitsGenes$symbol)
coxZScores = lapply(geneOfInterest, survivalAnalysisV2, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = FALSE,
                    survivalData = FALSE,
                    writeDirectory = '/microarray20200604_ewingSarcomaAllSets',
                    printData = FALSE)
esHitsGenes$coxz = unlist(coxZScores)
```

<div style="margin-bottom:50px;"></div>

Now we can make a plot for these scores.

```{r}
##########################################################################################
esHitsPlot = esHitsGenes %>%
  arrange(coxz) %>%
  mutate(index = seq(1,nrow(esHitsGenes),1))
esHitsPlot$pText = ifelse(esHitsPlot$symbol == 'YBX1', esHitsPlot$symbol, '')
esHitsPlot$pSize = ifelse(esHitsPlot$symbol == 'YBX1', 3, 1)
esHitsPlot$pColors = ifelse(esHitsPlot$symbol == 'YBX1', brewer.pal(8,'Spectral')[1], brewer.pal(8,'Greys')[5])
#esHitsPlot$pText = ifelse(esHitsPlot$coxz >= 2.5, esHitsPlot$symbol, '')
##
ggplot(esHitsPlot, aes(coxz, index)) +
  geom_point(size = esHitsPlot$pSize, color = esHitsPlot$pColors) +
  geom_text_repel(label = esHitsPlot$pText, size = 3, segment.size = 0.25) +
  labs(x = 'Cox Univariate Z Score', y = 'Gene index', title = 'Overall survival') +
  theme_classic()
ggsave(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/scatter_postelVinayOverallSurvivalTop250.pdf', sep = ''), 
       width = 3, height = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Extract the YB-1 data.

```{r}
##########################################################################################
expressionData = postelExpression
phenotypeData = postelPhenotype
geneOfInterest = 'YBX1'
coxZScores = lapply(geneOfInterest, survivalAnalysisV2, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = TRUE,
                    survivalData = TRUE,
                    writeDirectory = '/microarray20200604_ewingSarcomaAllSets',
                    printData = TRUE)
```

<div style="margin-bottom:50px;"></div>

This is the original version of the script where we just use quantiles for the cutoff.

```{r, message = FALSE}
##########################################################################################
expressionData = postelExpression
phenotypeData = postelPhenotype
geneOfInterest = as.list(esHitsGenes$symbol)
coxZScores = lapply(geneOfInterest, survivalAnalysis, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = FALSE,
                    survivalData = FALSE,
                    writeDirectory = '/microarray20200604_ewingSarcomaAllSets',
                    printData = FALSE)
esHitsGenes$coxz = unlist(coxZScores)
```

<div style="margin-bottom:50px;"></div>

Now we can make a plot for these scores.

```{r}
##########################################################################################
esHitsPlot = esHitsGenes %>%
  arrange(coxz) %>%
  mutate(index = seq(1,nrow(esHitsGenes),1))
esHitsPlot$pText = ifelse(esHitsPlot$symbol == 'YBX1', esHitsPlot$symbol, '')
esHitsPlot$pSize = ifelse(esHitsPlot$symbol == 'YBX1', 3, 1)
esHitsPlot$pColors = ifelse(esHitsPlot$symbol == 'YBX1', brewer.pal(8,'Spectral')[1], brewer.pal(8,'Greys')[5])
#esHitsPlot$pText = ifelse(esHitsPlot$coxz >= 2.5, esHitsPlot$symbol, '')
##
ggplot(esHitsPlot, aes(coxz, index)) +
  geom_point(size = esHitsPlot$pSize, color = esHitsPlot$pColors) +
  geom_text_repel(label = esHitsPlot$pText, size = 3, segment.size = 0.25) +
  labs(x = 'Cox Univariate Z Score', y = 'Gene index', title = 'Overall survival') +
  theme_classic()
ggsave(paste(baseRepository, '/microarray20200604_ewingSarcomaAllSets/scatter_postelVinayOverallSurvivalTop250Version1.pdf', sep = ''), 
       width = 3, height = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Extract the YB-1 data.

```{r}
##########################################################################################
expressionData = postelExpression
phenotypeData = postelPhenotype
geneOfInterest = 'YBX1'
coxZScores = lapply(geneOfInterest, survivalAnalysis, 
                    expressionData = expressionData, 
                    phenotypeData = phenotypeData, 
                    survivalPlot = TRUE,
                    survivalData = TRUE,
                    writeDirectory = '/microarray20200604_ewingSarcomaAllSets',
                    printData = TRUE)
```

## Session info

```{r}
sessionInfo()
```