---
title: "Analysis of published osteosarcoma microarray data"
output:
  html_notebook:
      code_folding: none
---

This document describes some analysis of some published microarray data from the study PMID: 21372215. The data will be obtained from ArrayExpress under the accession GSE21257.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('GEOquery')
library('illuminaHumanv2.db')
library('lumi')
library('annotate')
library('lumiHumanIDMapping')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

We need to call a source file that has some functions we will need. For simplicity and ease of tracking, I make a copy of this file in each directory that I will work in in order to be able to trace the exact one functions that are used, in case I change them and mistakenly don't push it to GitHub.

```{r}
source(paste(baseWorkspace, '/microarray20200603_osteosarcomaKuijjer/userDefinedFunctions.R', sep = ''))
```

## Data processing

So the first thing we will do is grab the data from arrayexpress.

```{r}
##################################################################################
osArrayData = getGEO('GSE21257',
             destdir = paste(baseRepository, '/microarray20200603_osteosarcomaKuijjer/', sep = ''),
             GSEMatrix = TRUE)
osData = osArrayData[[1]]
```
We now have an expression set and we can look at some of the features to see that we have all the information we need.

```{r}
##################################################################################
head(pData(osData))
head(exprs(osData))
```

The phenotype data looks good and has the information we need. The expression data looks ok, but the probe identifier is one I am not familiar with. On ArrayExpress they call them nuIDs: "Illumina human-6 v2.0 expression beadchip (using nuIDs as identifier)". It seems nuID is basically a compact identifier for the exact sequence of the probe ([some good info here](https://www.bioconductor.org/packages//2.7/bioc/vignettes/lumi/inst/doc/IlluminaAnnotation.pdf)). I basically follow that document for the below annotation. Accoding to the GEO site, the beadchip was Illumina human-6 v2.0 expression beadchip, which maps to "GPL6102", or illuminaHumanv2 R package.

```{r}
##################################################################################
osExpression = as_tibble(exprs(osData)) %>%
  mutate(nuID = row.names(exprs(osData)))
nuIDs = osExpression$nuID
osAnnotated = as_tibble(nuID2RefSeqID(nuIDs, lib.mapping='lumiHumanIDMapping', returnAllInfo = TRUE)) %>%
  mutate(nuID = nuIDs) %>%
  mutate(symbol = as.character(Symbol)) %>%
  mutate(entrez = as.character(EntrezID)) %>%
  dplyr::select(nuID, Accession, symbol) %>%
  left_join(osExpression)
saveRDS(osAnnotated, paste(baseRepository, '/microarray20200603_osteosarcomaKuijjer/dataset_osteoExpressionData.rds', sep = ''))
```
Ok, this looks good. Now lets get the phenotype data in good shape. The columns I want are geo_accession, characteristics_ch1, characteristics_ch1.1, characteristics_ch1.2, characteristics_ch1.3, characteristics_ch1.4, characteristics_ch1.5, characteristics_ch1.6.

```{r}
##################################################################################
osPheno = as_tibble(pData(osData)) %>%
  dplyr::select(geo_accession, characteristics_ch1:characteristics_ch1.6)
saveRDS(osPheno, paste(baseRepository, '/microarray20200603_osteosarcomaKuijjer/dataset_osteoPhenotypeData.rds', sep = ''))
```

Now we are all set to do the survival analysis. I will do this in another notebook.

NOTE - after some investigation, it appears the mapping from the lumi package was incorrect. Let's try this again. We can try using the feature data from the actual arrayExpress set.

```{r}
##################################################################################
osFeatures = as_tibble(fData(osData))
colnames(osFeatures) = c('nuID','illuminaID','illuminaGene','searchKey','genbank','symbol','sequence')
```

Now combine with the osExpression data set.

```{r}
##################################################################################
osExpression = as_tibble(exprs(osData)) %>%
  mutate(nuID = row.names(exprs(osData))) %>%
  left_join(osFeatures) %>%
  dplyr::select(nuID,genbank,symbol,GSM530667:GSM531352)
saveRDS(osExpression, paste(baseRepository, '/microarray20200603_osteosarcomaKuijjer/dataset_osteoExpressionData.rds', sep = ''))
```

YBX1 seems to be assigned correctly here, so we can move on.

## Session info

```{r}
sessionInfo()
```
