---
title: "KMT2D data"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of KMT2D data from Alessia.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message=FALSE, warning=FALSE}
##########################################################################################
#base packages
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
#generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
#baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToOthers'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToOthers'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Process the data into individual sets per replicate from the raw search output.

```{r}
##########################################################################################
samples = tibble('name' = c(rep('control',4),rep('kmt2d',4),rep('h3k4me1',4)),
                 'replicate' = c(rep(seq(1,4,1),3)))
##
samples$psmPath = file.path(baseRepository,
                            'proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond',
                            paste(samples$name,'_',samples$replicate,sep=''),
                            'psm.tsv')
samples$quantPath = file.path(baseRepository,
                              'proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/quants',
                              paste(samples$name,'_',samples$replicate,'_Matrix.txt',sep=''))
  
##
all(file.exists(samples$psmPath))
all(file.exists(samples$quantPath))

##
for (i in 1:nrow(samples)){
  quantTemp = read_tsv(samples$quantPath[i]) %>%
    dplyr::rename(scan = MS2ScanNumber,
                  area = ParentPeakArea) %>%
    dplyr::select(scan, area)
  ##
  pepTemp = read_tsv(samples$psmPath[i]) %>%
    dplyr::rename(accession = `Protein ID`,
                  gene = `Gene`,
                  sequence = `Peptide`,
                  mods = `Assigned Modifications`,
                  spectrum = Spectrum) %>%
    dplyr::select(accession, gene, sequence, mods, spectrum) %>%
    dplyr::filter(!grepl('^KRT', gene), !grepl('contam', accession)) %>%
    dplyr::mutate(scan = as.numeric(sub('.*[1-4]\\.(.*)\\.[0-9]+\\.[0-9]+','\\1',spectrum))) %>%
    dplyr::mutate(label = ifelse(grepl('R\\(10', mods), 'heavy',
                                 ifelse(grepl('K\\(8', mods), 'heavy', 'light'))) %>%
    dplyr::mutate(dataset = paste(samples[i,1],'_',samples[i,2],sep=''), numPsm = 1) %>%
    dplyr::select(scan, dataset, accession, gene, sequence, label, numPsm) %>%
    dplyr::mutate(label = factor(label, levels = c('light','heavy'))) %>%
    dplyr::left_join(quantTemp) %>%
    dplyr::group_by(dataset, accession, gene, sequence, label) %>%
    dplyr::summarise(numPsm = sum(numPsm, na.rm = TRUE), area = median(area, na.rm = TRUE)) %>%
    dplyr::mutate(numPeps = 1) %>%
    dplyr::group_by(dataset, accession, gene, label) %>%
    dplyr::summarise(numPsm = sum(numPsm, na.rm = TRUE), numPeps = sum(numPeps, na.rm = TRUE), area = median(area, na.rm = TRUE)) %>%
    tidyr::pivot_wider(id_cols = dataset:gene, names_from = 'label', values_from = c('numPsm','numPeps','area')) %>%
    dplyr::mutate(area_light = ifelse(is.na(area_light), 0, area_light)) %>%
    dplyr::mutate(area_heavy = ifelse(is.na(area_heavy), 0, area_heavy)) %>%
    dplyr::mutate(area_light_log2 = log2(area_light + 1000), area_heavy_log2 = log2(area_heavy + 1000)) %>%
    dplyr::mutate(log2HL = area_heavy_log2 - area_light_log2) %>%
    dplyr::rowwise() %>%
    dplyr::filter(sum(c(numPeps_light,numPeps_heavy), na.rm = TRUE) >= 2) %>%
    dplyr::arrange(-log2HL)
  ##
  saveRDS(pepTemp, paste(baseRepository,
                         '/proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/processedData/',
                         samples[i,1],
                         '_',
                         samples[i,2],
                         '_processed.rds',
                         sep = ''))
  ##
  write.table(pepTemp, paste(baseRepository,
                             '/proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/processedData/',
                             samples[i,1],
                             '_',
                             samples[i,2],
                             '_processed.csv',
                             sep = ''),
              quote = FALSE, col.names = TRUE, row.names = FALSE, sep = ',')
}
```

<div style="margin-bottom:50px;"></div>

Build a combined set with just the fold changes.

```{r}
##########################################################################################
samples$rdsFiles = file.path(baseRepository,
                            'proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/processedData',
                            paste(samples$name,'_',samples$replicate,'_processed.rds',sep=''))

##
allRdsFiles = tibble()
for (i in 1:nrow(samples)){
  inputTemp = readRDS(samples$rdsFiles[i]) %>%
    dplyr::select(dataset, accession, gene, log2HL)
  ##
  allRdsFiles = rbind(allRdsFiles, inputTemp)
}

##
allRdsFilesWide = allRdsFiles %>%
  tidyr::pivot_wider(id_cols = accession:gene, names_from = 'dataset', values_from = 'log2HL')

##
saveRDS(allRdsFilesWide,
        paste(baseRepository,
              '/proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/processedData/dataset_allSamplesLog2HL.rds', sep =''))
write.table(allRdsFilesWide,
        paste(baseRepository,
              '/proteomics20211208_alessiaKmt2dProcessedData/fragpipeAnalysisSecond/processedData/dataset_allSamplesLog2HL.csv', sep =''),
        quote = FALSE, col.names = TRUE, row.names = FALSE, sep = ',')
```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```
