---
title: "Analysis of survival data for Ewing sarcoma"
output:
  html_notebook:
      code_folding: none
---

I again want to look at survival data for Ewing sarcoma, but instead of just going set by set, I am going to combine all the sets I can find into one object and go from there. The details for each set will be presented as we go through.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('GEOquery')
library('org.Hs.eg.db')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

The first set is Savola et al. from GSE17679 that was published in PMID: 22084725. There appears to be two sets of replicates here, with a bunch of other cell line and other tissue samples analyzed as well that may need filtering downstream. First we will get the data.

```{r}
##########################################################################################
esArrayData = getGEO('GSE17679',
             destdir = paste(baseRepository, '/microarray20201203_dlg2SavolaSurvival/', sep = ''),
             GSEMatrix = TRUE)
esData = esArrayData[[1]]
```

<div style="margin-bottom:50px;"></div>

We now have an expression set and we can look at some of the features to see that we have all the information we need.

```{r}
##################################################################################
head(pData(esData))
head(exprs(esData))
head(fData(esData))
```

<div style="margin-bottom:50px;"></div>

First I just want to annotate the expression data a bit to include gene symbols, as these are what I will use for the survival analysis downstream. We can use the feature data to get this annotation.

```{r}
##################################################################################
esExpression = as_tibble(exprs(esData)) %>%
  mutate(ID = row.names(exprs(esData)))
esExpressionAnnotated = as_tibble(fData(esData)) %>%
  mutate(arraySymbolOne = sapply(strsplit(`Gene Symbol`, ' '), '[', 1)) %>%
  mutate(arraySymbolTwo = sapply(strsplit(`Gene Symbol`, ' '), '[', 3)) %>%
  mutate(arrayGenbank = sapply(strsplit(`RefSeq Transcript ID`, ' '), '[', 1)) %>%
  dplyr::select(ID, GB_ACC, arraySymbolOne, arraySymbolTwo, arrayGenbank) %>%
  filter(!is.na(arrayGenbank)) %>%
  left_join(esExpression)
```

<div style="margin-bottom:50px;"></div>

Convert the genbank ID's into gene symbols for use later on.

```{r}
##################################################################################
txdb = org.Hs.eg.db
geneKeys = unique(esExpressionAnnotated$arrayGenbank)
mappedGenes = AnnotationDbi::select(txdb, geneKeys, c('ACCNUM','SYMBOL'), 'ACCNUM')
colnames(mappedGenes) = c('arrayGenbank', 'symbol')
esExpressionAnnotatedTxdb = left_join(mappedGenes,esExpressionAnnotated) %>%
  dplyr::select(ID, symbol, arrayGenbank, arraySymbolOne, arraySymbolTwo, GSM439886:GSM441061)
```

<div style="margin-bottom:50px;"></div>

Now I want to look at the phenotype data. I think there are a bunch of samples here we don't want, so we should get rid of these. There appears to be PNET cases in here, cell lines, and some other tissues. I will keep only things classified as Ewing.

```{r}
##################################################################################
esPhenotype = as_tibble(pData(esData)) %>%
  dplyr::select(geo_accession, characteristics_ch1.1:characteristics_ch1.8) %>%
  filter(grepl('Ewing', characteristics_ch1.1)) %>%
  mutate(type = sub('state\\: (.*)$', '\\1', characteristics_ch1.3)) %>%
  mutate(age = sub('age\\: (.*)$', '\\1', characteristics_ch1.4)) %>%
  mutate(sex = sub('sex\\: (.*)$', '\\1', characteristics_ch1.5)) %>%
  mutate(efs = as.double(sub('efs \\(months\\)\\: (.*)$', '\\1', characteristics_ch1.6))) %>%
  mutate(ovs = as.double(sub('ovs \\(months\\)\\: (.*)$', '\\1', characteristics_ch1.7))) %>%
  mutate(status = sub('status\\: (.*)$', '\\1', characteristics_ch1.8)) %>%
  dplyr::select(geo_accession, type:status)
esPhenotype$status = ifelse(esPhenotype$status == 'Dead', 2, 1)
```

<div style="margin-bottom:50px;"></div>

Now we can save the two datasets.

```{r}
##################################################################################
saveRDS(esExpressionAnnotatedTxdb, paste(baseRepository, '/microarray20201203_dlg2SavolaSurvival/dataset_savolaExpression.rds', sep = ''))
saveRDS(esPhenotype, paste(baseRepository, '/microarray20201203_dlg2SavolaSurvival/dataset_savolaPhenotype.rds', sep = ''))
```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##################################################################################
sessionInfo()
```