---
title: "Analysis of FAIMS Pro CV Scans"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of MS data acquired across different CV values using the FAIMS Pro device on the Orbitrap Fusion MS. The sample was a tryptic digest of a HeLa cell lysate that was analyzed in both unlabeled and TMT labeled formats. The runs were all 60-minutes on a 150um ID x 20cm length column packed with 1.9um C18 beads, running at a flow rate of 0.8uL/min. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
###########################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

Set the base folders where we will do the work.

```{r}
###########################################################
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/faimsProTesting'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/faimsProTesting'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing - unlabeled samples

The first thing I want to look at is the charge state distribution across the different CV values. I will use the RawTools output files for this and only focus on things that triggered MS2 scans (regardless of whether they generated a peptide hit or not). 

Read the files into a list.

```{r, message = FALSE, warning = FALSE}
###########################################################
cvFiles = list.files(paste(baseRepository, '/nonTmtQuantFiles/', sep = ''), pattern = '_Matrix.txt', full.names = TRUE)
cvData = lapply(cvFiles, read_tsv)
names(cvData) = sub('.*400ng_(.*)\\.raw_Matrix\\.txt', '\\1', cvFiles)
```

<div style="margin-bottom:50px;"></div>

Process the data for each list entry to get the scan numbers for each charge.

```{r}
###########################################################
cvCharges = lapply(cvData, function(x) table(x$PrecursorCharge))
names(cvCharges) = names(cvData)
cvChargesDf = as.data.frame(do.call(rbind, cvCharges))
colnames(cvChargesDf) = c('z2','z3','z4')
cvChargesDf$sample = row.names(cvChargesDf)
row.names(cvChargesDf) = NULL
##
cvChargesTib = as_tibble(cvChargesDf) %>%
  filter(!grepl('_', sample)) %>%
  mutate(cv = sub('CV(.*)$', '\\1', sample)) %>%
  pivot_longer(cols = z2:z4, names_to = 'charge', values_to = 'numberOfPrecursors')
cvChargesTib$cv = factor(cvChargesTib$cv, levels = seq(110,10,-5))
```

<div style="margin-bottom:50px;"></div>

Make a plot of the non-TMT sample data.

```{r}
###########################################################
ggplot(cvChargesTib, aes(cv, numberOfPrecursors, group = charge, color = charge)) +
  geom_point(size = 2) +
  geom_line(size = 0.75) +
  labs(x = 'CV', y = 'Number of Precursors', title = 'Unlabeled charge state distribution') +
  scale_color_manual(values = brewer.pal(9,'Greys')[c(9,6,4)]) +
  scale_y_continuous(limits = c(0, 2e4), breaks = seq(0,2e4,2500)) +
  theme_classic() +
  theme(panel.grid.major = element_line(size = 0.25), 
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 6),
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
ggsave(paste(baseRepository, '/scatter_cvChargeDistributionNonTmt.pdf', sep = ''),
       height = 4, width = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

This looks pretty similar to what has been shown in other papers related to FAIMS Pro by the Coon/Gygi/Olsen labs. This plot actually does a pretty good job of highlighting why a single CV isn't that useful as you end up filtering a lot of potential peptide ions out. To see how this actually translates to identifications, I processed the data in MaxQuant.

Parse the identification data.

```{r, message = FALSE, warning = FALSE}
###########################################################
idData = read_tsv(paste(baseRepository, '/nonTmtMaxQuantResults/summary.txt', sep = '')) %>%
  dplyr::select(`Raw file`, `MS/MS`, `MS/MS Identified`, `MS/MS Identified [%]`) %>%
  mutate(cv = sub('.*_CV(.*)$', '\\1', `Raw file`)) %>%
  filter(!grepl('Total', `Raw file`))
##
idData$cv = factor(idData$cv, levels = seq(110,10,-5))
```

<div style="margin-bottom:50px;"></div>

Plot the identification data.

```{r}
###########################################################
ggplot(idData, aes(cv, `MS/MS`)) +
  geom_point(size = 2) +
  scale_y_continuous(limits = c(0, 3e4), breaks = seq(0,3e4,5000)) +
  labs(x = 'CV', y = 'Number of MS/MS', title = 'Unlabeled scan distribution') +
  theme_classic() +
  theme(panel.grid.major = element_line(size = 0.25), 
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 6),
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
```

<div style="margin-bottom:50px;"></div>

## Data processing - TMT labeled samples

Read the files into a list.

```{r, message = FALSE, warning = FALSE}
###########################################################
cvFiles = list.files(paste(baseRepository, '/tmtQuantFiles/', sep = ''), pattern = '_Matrix.txt', full.names = TRUE)
cvData = lapply(cvFiles, read_tsv)
names(cvData) = sub('.*400ng_(.*)\\.raw_Matrix\\.txt', '\\1', cvFiles)
```

<div style="margin-bottom:50px;"></div>

Process the data for each list entry to get the scan numbers for each charge.

```{r}
###########################################################
cvCharges = lapply(cvData, function(x) table(x$PrecursorCharge))
names(cvCharges) = names(cvData)
cvChargesDf = as.data.frame(do.call(rbind, cvCharges))
colnames(cvChargesDf) = c('z2','z3','z4')
cvChargesDf$sample = row.names(cvChargesDf)
row.names(cvChargesDf) = NULL
##
cvChargesTib = as_tibble(cvChargesDf) %>%
  filter(!grepl('_', sample)) %>%
  mutate(cv = sub('CV(.*)$', '\\1', sample)) %>%
  pivot_longer(cols = z2:z4, names_to = 'charge', values_to = 'numberOfPrecursors')
cvChargesTib$cv = factor(cvChargesTib$cv, levels = seq(110,10,-5))
```

<div style="margin-bottom:50px;"></div>

Make a plot of the non-TMT sample data.

```{r}
###########################################################
ggplot(cvChargesTib, aes(cv, numberOfPrecursors, group = charge, color = charge)) +
  geom_point(size = 2) +
  geom_line(size = 0.75) +
  labs(x = 'CV', y = 'Number of Precursors', title = 'TMT labeled charge state distribution') +
  scale_color_manual(values = brewer.pal(9,'Greys')[c(9,6,4)]) +
  scale_y_continuous(limits = c(0, 2e4), breaks = seq(0,2e4,2500)) +
  theme_classic() +
  theme(panel.grid.major = element_line(size = 0.25), 
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 6),
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
ggsave(paste(baseRepository, '/scatter_cvChargeDistributionTmt.pdf', sep = ''),
       height = 4, width = 4, useDingbats = FALSE)
```







