---
title: "MS275 in NRF2 HiBit cells"
author: "Christopher Hughes"
format: html
---

## Details

This document describes processing of plate reader data measuring NRF2-HiBit levels in A673p1 cells in response to a variety of conditions.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| message: false
#| warning: false
##########################################################################################
library('tidyverse')
library('RColorBrewer')
library('ggrepel')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

## Data processing

Read in the plate reader data.

```{r}
##########################################################################################
##read in the data
hibit = read_tsv(paste(baseRepository, '/plateReader20230329_a673p1Nrf2ArseniteMs275/plateReader20230329_a673p1Nrf2ArseniteMs275Parsed.txt', sep = ''), show_col_types = FALSE) %>%
  dplyr::mutate(condition = factor(sub('.*_(.*)[1-6]','\\1',Sample), levels = c('dmso','ms275','ars','combo')))

##plot the data
hibitSummary = read_tsv(paste(baseRepository, '/plateReader20230329_a673p1Nrf2ArseniteMs275/plateReader20230329_a673p1Nrf2ArseniteMs275Parsed.txt', sep = ''), show_col_types = FALSE) %>%
  dplyr::mutate(condition = factor(sub('.*_(.*)[1-6]','\\1',Sample), levels = c('dmso','ms275','ars','combo'))) %>%
  dplyr::group_by(condition) %>%
  dplyr::summarise(mean = mean(Value, na.rm = TRUE), sd = sd(Value, na.rm = TRUE))

ggplot(hibitSummary, aes(condition, mean, fill = condition)) +
  geom_col(position = position_dodge(width = 1)) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,position=position_dodge(.9), color = brewer.pal(4,'Greys')[4]) +
  geom_point(data = hibit, aes(condition, Value), color = brewer.pal(4,'Greys')[4], position = position_jitter(width = 0.15), size = 0.5) +
  scale_fill_manual(values = c(brewer.pal(4,'GnBu')[4], rep(brewer.pal(4,'OrRd')[4],3))) +
  labs(x = 'Treatment', y = 'HiBit luminescence', title = 'NRF2 levels') +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/plateReader20230329_a673p1Nrf2ArseniteMs275/barplot_hibitNrf2Levels.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```