---
title: "Analysis of DLG2 siRNA and KRAB data"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of qPCR data from cells treated with siRNA against DLG2 as well as KRAB mediated CRISPR inhibition.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message=FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('ggrepel')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read in the qPCR results data and process the sirna samples.

```{r}
##########################################################################################
qpcrData = read_tsv(paste(baseRepository, '/qpcr20220330_dlg2KrabSamKoClones/qpcr20220330_dlg2KrabSamKoClones.txt', sep = ''))

##sirna data
qpcrDataSub = qpcrData %>%
  dplyr::filter(grepl('dsi',sampleName)) %>%
  dplyr::mutate(replicate = rep(rep(seq(1,3,1),each = 5),2)) %>%
  tidyr::pivot_wider(names_from = targetName, values_from = Ct, names_sep = "_") %>%
  dplyr::mutate(deltaCt = DLG2n3 - GAPDH)

##
qpcrNegCt = qpcrDataSub %>%
  dplyr::filter(grepl('Neg', sampleName))
qpcrNegCtValue = mean(qpcrNegCt$deltaCt)
  
##
qpcrSirna = qpcrDataSub %>%
  dplyr::filter(!grepl('Neg', sampleName)) %>%
  dplyr::mutate(deltaDeltaCt = deltaCt - qpcrNegCtValue) %>%
  dplyr::mutate(log2FC = log2(2^-deltaDeltaCt)) %>%
  dplyr::mutate(sampleName = factor(sampleName, levels = c('dsiPos','dsiDlg2_1','dsiDlg2_2','dsiDlg2_3')))

##
ggplot(qpcrSirna, aes(sampleName, log2FC, color = sampleName)) +
  geom_point(size = 2) +
  labs(x = 'Sample name', y = 'log2(Fold change change relative to negative)', title = 'DLG2 siRNA knockdown') +
  scale_color_manual(values = c(brewer.pal(3,'Greys')[2], rep(brewer.pal(3,'OrRd')[3],3))) +
  theme_classic() +
  theme(legend.position = 'none')
#
ggsave(paste(baseRepository, '/qpcr20220330_dlg2KrabSamKoClones/scatter_dlg2DsirnaKnockdown.pdf',sep=''),
       height=2,width=2,useDingbats=FALSE)
```

<div style="margin-bottom:50px;"></div>

Read in the qPCR results data and process the krab samples.

```{r}
##########################################################################################
qpcrData = read_tsv(paste(baseRepository, '/qpcr20220330_dlg2KrabSamKoClones/qpcr20220330_dlg2KrabSamKoClones.txt', sep = ''))

##sirna data
qpcrDataSub = qpcrData %>%
  dplyr::filter(grepl('krab',sampleName)) %>%
  dplyr::mutate(replicate = rep(rep(seq(1,3,1),each = 15),2)) %>%
  tidyr::pivot_wider(names_from = targetName, values_from = Ct, names_sep = "_") %>%
  dplyr::mutate(deltaCt = DLG2n3 - GAPDH)

##
qpcrNegCt = qpcrDataSub %>%
  dplyr::filter(grepl('Neg', sampleName))
qpcrNegCtValue = mean(qpcrNegCt$deltaCt)
  
##
qpcrSirna = qpcrDataSub %>%
  dplyr::filter(!grepl('Neg', sampleName)) %>%
  dplyr::mutate(deltaDeltaCt = deltaCt - qpcrNegCtValue) %>%
  dplyr::mutate(log2FC = log2(2^-deltaDeltaCt)) %>%
  dplyr::mutate(sampleName = factor(sampleName, levels = sampleName[1:12]))

##
ggplot(qpcrSirna, aes(sampleName, log2FC, color = sampleName)) +
  geom_point(size = 2) +
  labs(x = 'Sample name', y = 'log2(Fold change relative to negative)', title = 'DLG2 KRAB inhibition') +
  scale_color_manual(values = c(rep(brewer.pal(3,'OrRd')[3],12))) +
  theme_classic() +
  theme(legend.position = 'none')
#
ggsave(paste(baseRepository, '/qpcr20220330_dlg2KrabSamKoClones/scatter_dlg2KrabInhibition.pdf',sep=''),
       height=2,width=2,useDingbats=FALSE)
```


