---
title: "ChIPseq data analysis from Riggi"
output:
  html_notebook:
      code_folding: none
---

This document details working with aligned ChIPseq data from the Riggi paper, looking for DLG2 patterns.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning = FALSE}
##########################################################################################
library('tidyverse')
library('GenomicRanges')
library('ggplot2')
library('wiggleplotr')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

First get the gtf file we used for alignment and will use for annotation below.

```{r}
##########################################################################################
totalGtf = as.data.frame(rtracklayer::import.gff('C:/Users/chughes/Documents/bccrc/databases/hg38/refseq_gtf/GRCh38_latest_genomic_feb2021.gtf.gz'))
```

<div style="margin-bottom:50px;"></div>

We want to look at H3K4me3 signal in the region of the potential new DLG2 start site. So, we will look through these data for both cell lines. I ended up modifying the plots generated below in Illustrator after the fact just to make the data a bit nicer to visualize. If you want to have the gene actually plotted, you need to not output the plots to the list using the 'return_subplots_list' trigger in the plotCoverage command.

```{r}
##########################################################################################
transOfInterest = c('NM_001377978.1')
dlg2Gtf = totalGtf %>%
  filter(transcript_id %in% transOfInterest) %>%
  dplyr::rename(gene_name = gene_id, gene_id = seqnames) %>%
  mutate('seqnames' = 'chr11')
##
dlg2Meta = tibble('transcript_id' = unique(dlg2Gtf$transcript_id),
                  'gene_id' = unique(dlg2Gtf$gene_id),
                  'gene_name' = 'DLG2',
                  'strand' = '-')
##
dlg2Exons = makeGRangesListFromDataFrame(dlg2Gtf %>%
  filter(grepl('exon', type)), split.field = 'transcript_id')
##
dlg2Cds = makeGRangesListFromDataFrame(dlg2Gtf %>%
  filter(grepl('CDS', type)), split.field = 'transcript_id')
##
sampleData = tibble(sample_id = factor(c('a673a','a673b','a673c','a673d','a673e')), 
                    condition = factor(c('h3k4me3','h3k27ac','h3k4me1','h3k27me3','wce'), levels = c('h3k4me3','h3k27ac','h3k4me1','h3k27me3','wce')),
                    scaling_factor = 1)
##
bigwigFiles = c(paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593989.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593986.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593987.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593988.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593991.chr11.bw', sep = ''))
##
sampleData = sampleData %>%
  dplyr::mutate(bigWig = bigwigFiles)
as.data.frame(sampleData)
##
#trackData = dplyr::mutate(sampleData, track_id = condition, colour_group = condition)
trackData = dplyr::mutate(sampleData, track_id = condition, colour_group = sample_id)
##
covPlot = plotCoverage(dlg2Exons, dlg2Cds, 
             dlg2Meta, trackData,
             heights = c(2,0.5),
             rescale_introns = FALSE, 
             fill_palette = c(rep(brewer.pal(8,'Spectral')[1],4), brewer.pal(8,'Spectral')[8]),
             region_coords = c(83626956, 83760000),
             coverage_type = 'both',
             return_subplots_list = TRUE)
##
covPlot$coverage_plot +
  scale_y_continuous(limits = c(0,4), breaks = seq(0,6,1))
##
ggsave(paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/coverage_dlg2RefseqChipSeqHistonesA673.pdf', sep = ''),
         height = 4, width = 4, useDingbats = FALSE)



#############################################
sampleData = tibble(sample_id = factor(c('sknmca','sknmcb','sknmcc','sknmcd','sknmce')), 
                    condition = factor(c('h3k4me3','h3k27ac','h3k4me1','h3k27me3','wce'), levels = c('h3k4me3','h3k27ac','h3k4me1','h3k27me3','wce')),
                    scaling_factor = 1)
##
bigwigFiles = c(paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593964.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593961.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593962.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593963.chr11.bw', sep = ''),
                paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/SRR1593966.chr11.bw', sep = ''))
##
sampleData = sampleData %>%
  dplyr::mutate(bigWig = bigwigFiles)
as.data.frame(sampleData)
##
#trackData = dplyr::mutate(sampleData, track_id = condition, colour_group = condition)
trackData = dplyr::mutate(sampleData, track_id = condition, colour_group = sample_id)
##
covPlot = plotCoverage(dlg2Exons, dlg2Cds, 
             dlg2Meta, trackData,
             heights = c(2,0.5),
             rescale_introns = FALSE, 
             fill_palette = c(rep(brewer.pal(8,'Spectral')[1],4), brewer.pal(8,'Spectral')[8]),
             region_coords = c(83626956, 83760000),
             coverage_type = 'both',
             return_subplots_list = TRUE)
##
covPlot$coverage_plot +
  scale_y_continuous(limits = c(0,2), breaks = seq(0,6,1))
##
ggsave(paste(baseRepository, '/sequencing20201208_ewsFil1RnaSeqRiggiPmid25453903/chipSeq/coverage_dlg2RefseqChipSeqHistonesSknmc.pdf', sep = ''),
         height = 4, width = 4, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```



