---
title: "Analysis of Sonya translatome data"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of proteomics data acquired to look at the translatome of cells from Sonya. First, we ran the samples with and without targeted mass difference to see which worked better. This document is for comparing these two data sets.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToOthers'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToOthers'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read the protein annotation details.

```{r}
##########################################################################################
proteinAnnotation = readRDS(paste(baseRepository, '/proteomics20201110_translationSonya/uniprotHumanNov2020.fasta.annotated.rds', sep = ''))
```

<div style="margin-bottom:50px;"></div>

Read in the psm data and process into a peptide set.

```{r}
##########################################################################################
psm = read_tsv(paste(baseRepository, '/proteomics20201110_translationSonya/20201106_SonyaTranslationSet1_silac10R8K-TMT10_HpH_Default_PSM_Report.txt', sep = '')) %>%
  dplyr::select(`Protein(s)`, Sequence, `Modified Sequence`, `Spectrum File`, `Spectrum Title`, `Variable Modifications`) %>%
  mutate(fraction = factor(sub('.*HpH_(.*)\\.raw$', '\\1', `Spectrum File`), levels = seq(1,36,1))) %>%
  mutate(accession = `Protein(s)`) %>%
  mutate(scan = as.numeric(sub('Spectrum_(.*)', '\\1', `Spectrum Title`))) %>%
  mutate(sequence = Sequence) %>%
  mutate(modSequence = `Modified Sequence`) %>%
  mutate(varMods = `Variable Modifications`) %>%
  left_join(proteinAnnotation) %>%
  dplyr::select(fraction, scan, accession, gene, detectablePeptides, sequence, modSequence, varMods)
```

<div style="margin-bottom:50px;"></div>

Now we need the quantification data. I will use a function to process these data.

```{r}
#######################################################################
#######################################################################
##this function goes through the provided quant files and returns
##a parsed object for each. It returns signal to noise ratio for quant
#######################################################################
combineQuantFiles = function(filePath, ...){
  quantData = read_tsv(filePath) %>%
    dplyr::select(MS2ScanNumber, `126Intensity`:`131CIntensity`)
  colnames(quantData) = c('scan','tmt10plex_126','tmt10plex_127N','tmt10plex_127C','tmt10plex_128N',
                  'tmt10plex_128C','tmt10plex_129N','tmt10plex_129C','tmt10plex_130N','tmt10plex_130C','tmt10plex_131N','tmt10plex_131C')
  ##
  fraction = sub('.*HpH_(.*)\\.raw_Matrix\\.txt$', '\\1', filePath)
  quantData$fraction = fraction
  print(paste('Processing file for fraction ', fraction, '.', sep = ''))
  ##
  return(quantData)
}
```

<div style="margin-bottom:50px;"></div>

We can use this function to parse the quant data into a single object.

```{r, message = FALSE, warning = FALSE}
##########################################################################################
quantFiles = as.list(list.files(paste(baseRepository, '/proteomics20201110_translationSonya/quantFiles/', sep = ''),
           pattern = '_Matrix.txt', full.names = TRUE))
##
quantDataSet = lapply(quantFiles, combineQuantFiles)
```

<div style="margin-bottom:50px;"></div>

Now combine the quant data into a single data frame and with the previously processed PSM data.

```{r}
##########################################################################################
allQuantData = do.call('rbind', quantDataSet)
psmQuant = psm %>%
  left_join(allQuantData)
```

<div style="margin-bottom:50px;"></div>

Add light and heavy annotation.

```{r}
##########################################################################################
psmQuant$label = ifelse(grepl('Arginine', psmQuant$varMods), 'heavy',
                        ifelse(grepl('K\\+8', psmQuant$varMods), 'heavy', 'light'))
##
table(psmQuant$label)
```

<div style="margin-bottom:50px;"></div>

We detect so few heavy peptides. That is concerning. It means we aren't really getting a great snapshot of new protein synthesis. Maybe the TMD data will be better here?

## Session info

```{r}
##########################################################################################
sessionInfo()
```






