---
title: "EwS microarray data"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of microarray data from various EwS datasets.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message=FALSE, warning=FALSE}
##########################################################################################
#base packages
library('tidyverse')
library('ggplot2')
library('RColorBrewer')

#microarray specific packages
#library('clariomdhumantranscriptcluster.db')
library('oligo')
library('ff')
library('affycoretools')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
#generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
#baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
```

<div style="margin-bottom:50px;"></div>

## Data processing

The first thing I need to do is set the annotation. I use a CDF from BrainArray (in this case, version 25). The commands below are to build the libraries I need for the different arrays used. You only need to do the package install once, and then just load the library from then on. 

```{r}
##########################################################################################
#these are the install commands for my windows machine
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133plus2/pd.hgu133plus2.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133plus2/hgu133plus2hsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')

#these were for when I had to run it on linux because of memory issues
#install.packages('/home/chughes/databases/microarray/clariomDHuman/pd.clariomdhuman.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('/home/chughes/databases/microarray/clariomDHuman/clariomdhumanhsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')

#####################
##library for clariomd
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/clariomd/pd.clariomdhuman.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/clariomd/clariomdhumanhsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.clariomdhuman.hs.refseq')
library('clariomdhumanhsrefseq.db')

##library for hgu133a
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133a/pd.hgu133a.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133a/hgu133ahsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.hgu133a.hs.refseq')
library('hgu133ahsrefseq.db')

##library for hgu133a2
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133a2/pd.hgu133a2.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133a2/hgu133a2hsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.hgu133a2.hs.refseq')
library('hgu133a2hsrefseq.db')

##library for hgu133plus2
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133plus2/pd.hgu133plus2.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hgu133plus2/hgu133plus2hsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.hgu133plus2.hs.refseq')
library('hgu133plus2hsrefseq.db')

##library for huex10st
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/huex10st/pd.huex10st.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/huex10st/huex10sthsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.huex10st.hs.refseq')
library('huex10sthsrefseq.db')

##library for hugene10st
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hugene10st/pd.hugene10st.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hugene10st/hugene10sthsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.hugene10st.hs.refseq')
library('hugene10sthsrefseq.db')

##library for hugene20st
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hugene20st/pd.hugene20st.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/hugene20st/hugene20sthsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.hugene20st.hs.refseq')
library('hugene20sthsrefseq.db')

##library for u133aaofav2
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/u133aaofav2/pd.u133aaofav2.hs.refseq_25.0.0.tar.gz', repos = NULL, type = 'source')
#install.packages('C:/Users/chris/OneDrive/Documents/bccrc/databases/microarrays/u133aaofav2/u133aaofav2hsrefseq.db_25.0.0.tar.gz', repos = NULL, type = 'source')
library('pd.u133aaofav2.hs.refseq')
library('u133aaofav2hsrefseq.db')
```

<div style="margin-bottom:50px;"></div>

Read the annotation and dataset table.

```{r}
##########################################################################################
samples = read_tsv(paste(baseRepository,'/microarray20211209_ewsExpressionDatasets/samples.txt',sep=''))
```

<div style="margin-bottom:50px;"></div>

Work through the different accessions and save the output data for later use. 

```{r}
##########################################################################################
#following directions here: https://support.bioconductor.org/p/72378/#72552
#list.celfiles(paste(baseRepository, '/microarray20211209_ewsExpressionDatasets/GSE6930', sep = ''))

for (i in 2:nrow(samples)){
  print(paste('working on dataset', samples[i,1]))
  print(paste('reference package being used is: pd.',samples[i,2],'.hs.refseq',sep=''))
  print(paste('annotation package being used is: ',samples[i,2],'hsrefseq.db',sep=''))
  ##
  dat = read.celfiles(list.celfiles(paste(baseRepository, 
                                          '/microarray20211209_ewsExpressionDatasets/', 
                                          samples[i,1], 
                                          sep = ''), 
                                    full.names=TRUE), 
                      pkgname = paste('pd.',samples[i,2],'.hs.refseq',sep=''))
  ##
  rmaEset = rma(dat)
  annotatedEset = annotateEset(rmaEset, paste(samples[i,2],'hsrefseq.db',sep=''))
  expressionData = as_tibble(exprs(annotatedEset)) %>%
    dplyr::mutate(PROBEID = rownames(annotatedEset))
    #dplyr::rename_with(~ sub('(.*)\\.CEL', '\\1', .x))
  annotatedEdata = as_tibble(fData(annotatedEset)) %>%
    left_join(expressionData)
  ##
  saveRDS(annotatedEdata, 
          paste(baseRepository,'/microarray20211209_ewsExpressionDatasets/', samples[i,1],'/',samples[i,1],'.processed.rds', sep = ''))
  ##
  #clean out previous result
  rm(dat)
  rm(rmaEset)
  rm(annotatedEset)
  rm(expressionData)
  rm(annotatedEdata)
}
```




















<div style="margin-bottom:50px;"></div>

Now read in the data files and process them with rma from the oligo package. 

```{r}
##########################################################################################
#following directions here: https://support.bioconductor.org/p/72378/#72552
list.celfiles(paste(baseRepository, '/microarray20211209_ewsExpressionDatasets/', sep = ''))
dat = read.celfiles(list.celfiles(paste(baseRepository, '/microarray20211209_ewsExpressionDatasets', sep = ''), full.names=TRUE)[1:3], pkgname = 'pd.hgu133plus2.hs.refseq')

rmaEset = rma(dat)
annotatedEset = annotateEset(rmaEset, hgu133plus2hsrefseq.db)
expressionData = as_tibble(exprs(annotatedEset)) %>%
  dplyr::mutate(PROBEID = rownames(annotatedEset)) %>%
  dplyr::rename_with(~ sub('.*_(.*)\\.CEL', '\\1', .x))
annotatedEdata = as_tibble(fData(annotatedEset)) %>%
  left_join(expressionData)





allData = tibble()
for (i in seq(1,108,27)){
  dat = read.celfiles(list.celfiles(paste(baseRepository, '/microarray20211209_ewsExpressionDatasets', sep = ''), full.names=TRUE)[i:i+26], pkgname = 'pd.clariomdhuman.hs.refseq')
  rmaEset = rma(dat)
  annotatedEset = annotateEset(eset, clariomdhumanhsrefseq.db)
  expressionData = as_tibble(exprs(annotatedEset)) %>%
    dplyr::mutate(PROBEID = rownames(annotatedEset)) %>%
    dplyr::rename_with(~ sub('(.*)_[0-9]+\\-[0-9]+\\.CEL', '\\1', .x))
  annotationData = as_tibble(fData(annotatedEset)) %>%
    left_join(expressionData)
  #
  allData = allData %>%
    left_join(annotationData)
}


```



