---
title: "total proteome"
author: "Christopher Hughes"
format: html
---

## Details

This document describes processing of the total proteome data from A673p1 cells after treatment with MS275 and arsenite.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| message: false
#| warning: false
##########################################################################################
library('tidyverse')
library('iq')
library('DEqMS')
library('matrixStats')
library('RColorBrewer')
library('ggrepel')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

## Data processing

Read in data from DIA-NN and process with the IQ package.

```{r}
##########################################################################################
##first we can filter and parse the report data into a better format for IQ
raw = read_tsv('D:/msDataTemp/proteomics20230223_a673p1Ybx1HibitTotalProteome/report.tsv', show_col_types = FALSE) %>%
  dplyr::mutate(File.Name = sub('.*totalProteome_(.*)_dia_[123]\\.mzML', '\\1', File.Name)) %>%
  dplyr::filter(Q.Value <= 0.01)

##save the data
write.table(raw, 'D:/msDataTemp/proteomics20230223_a673p1Ybx1HibitTotalProteome/reportParsed.tsv', 
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')


##run the IQ analysis
iqResult = process_long_format('D:/msDataTemp/proteomics20230223_a673p1Ybx1HibitTotalProteome/reportParsed.tsv', 
                           output_filename = 'D:/msDataTemp/proteomics20230223_a673p1Ybx1HibitTotalProteome/reportParsedIqQuant.tsv',
                           annotation_col = 'Genes',
                           normalization = 'none',
                           peptide_extractor = function(x) gsub("[0-9].*$", "", x))

##parse the IQ result
iqQuantParsed = read_tsv('D:/msDataTemp/proteomics20230223_a673p1Ybx1HibitTotalProteome/reportParsedIqQuant.tsv', show_col_types = FALSE) %>%
  dplyr::rename(accession = 'Protein.Group',
                symbol = 'Genes') %>%
  dplyr::select(accession:n_peptides, 
                dmso_rep1, dmso_rep2, dmso_rep3,
                ms275_rep1, ms275_rep2, ms275_rep3,
                ars_rep1, ars_rep2, ars_rep3,
                combo_rep1, combo_rep2, combo_rep3) %>%
  dplyr::filter(!grepl(';', accession) & !is.na(symbol))
#  dplyr::filter(rowSums(is.na(dmso_rep1:dmso_rep3)) == 0 |
#                  rowSums(is.na(ms275_rep1:ms275_rep3)) == 0 |
#                  rowSums(is.na(ars_rep1:ars_rep3)) == 0 |
#                  rowSums(is.na(combo_rep1:combo_rep3)) == 0)

##save the data
saveRDS(iqQuantParsed, paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_processedTotalProteomeData.rds', sep = ''))
write.table(iqQuantParsed, paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_processedTotalProteomeData.tsv', sep = ''), 
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')
```

Perform DEqMS analysis.

```{r}
##########################################################################################
##extract the data we need into a suitable data frame
##I decided to do this comparison one at a time so I can make some individual adjustments and because the MS275 didn't really work well (the treatment)
conditionsToCompare = c('ms275','ars','combo')
for (i in 1:3){
  dat = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_processedTotalProteomeData.rds', sep = ''))) %>%
    #dplyr::select(-accession, -symbol, -n_fragments, -n_peptides) %>%
    dplyr::select(accession, symbol, 
                  dmso_rep1:dmso_rep3, 
                  paste(conditionsToCompare[i], '_rep1', sep=''),
                  paste(conditionsToCompare[i], '_rep2', sep=''),
                  paste(conditionsToCompare[i], '_rep3', sep=''))
  dat$samples123 = rowSums(is.na(dat[,3:5]))
  dat$samples456 = rowSums(is.na(dat[,6:8]))
  colnames(dat)[6:8] = c('treat_rep1','treat_rep2','treat_rep3')
  #
  datFiltered = dat %>%
    dplyr::filter(samples123 < 1 | samples456 < 1) %>%
    dplyr::mutate(dmso_rep1 = ifelse(samples123 == 3, min(dmso_rep1, na.rm = TRUE), dmso_rep1),
                  dmso_rep2 = ifelse(samples123 == 3, min(dmso_rep2, na.rm = TRUE), dmso_rep2),
                  dmso_rep3 = ifelse(samples123 == 3, min(dmso_rep3, na.rm = TRUE), dmso_rep3)) %>%
    dplyr::mutate(treat_rep1 = ifelse(samples456 == 3, min(treat_rep1, na.rm = TRUE), treat_rep1),
                  treat_rep2 = ifelse(samples456 == 3, min(treat_rep2, na.rm = TRUE), treat_rep2),
                  treat_rep3 = ifelse(samples456 == 3, min(treat_rep3, na.rm = TRUE), treat_rep3)) %>%
    dplyr::select(-samples123, -samples456)
  #
  rownames(datFiltered) = datFiltered$accession
  dat.log = datFiltered %>%
    dplyr::select(-accession, -symbol)
  #dat.log = na.omit(datFiltered)
  #boxplot(dat.log, las = 2, main = 'total protein abundance')
  
  ##make a design table
  cond = as.factor(c(rep('dmso',3),
                    rep('treat',3)))
  design = model.matrix(~0+cond)
  colnames(design) = gsub('cond', '', colnames(design))
  
  ##make the contrasts
  x = c('treat-dmso')
  contrast = makeContrasts(contrasts = x, levels = design)
  fit1 = lmFit(dat.log, design)
  fit2 = contrasts.fit(fit1, contrasts = contrast)
  fit3 = eBayes(fit2)

  ##add information about PSM counts
  forPsm = readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_processedTotalProteomeData.rds', sep = ''))
  psm.count.table = data.frame(count = forPsm$n_peptides,
                               row.names = forPsm$accession)
  fit3$count = psm.count.table[rownames(fit3$coefficients),'count']
  fit4 = spectraCounteBayes(fit3)
  #VarianceBoxplot(fit4, n = 30, main = 'total proteome dataset', xlab = 'psm count')
  #VarianceScatterplot(fit4, main = 'total proteome dataset')
  
  ##extract the data
  #head(fit4$coefficients)
  toAnnotate = forPsm[,1:4]
  deqmsResults = tibble(outputResult(fit4, coef_col = 1)) %>%
    dplyr::rename(accession = 'gene') %>%
    dplyr::left_join(toAnnotate)
  #
  saveRDS(deqmsResults, paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_', conditionsToCompare[i], '-dmso.rds', sep = ''))
  write.table(deqmsResults, paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_', conditionsToCompare[i], '-dmso.tsv', sep = ''),
              col.names = TRUE, row.names = FALSE, sep = '\t', quote = FALSE)
}
```

Read in the comparison data and make a volcano plot.

```{r}
##########################################################################################
##read in the data for arsenite and plot
##assign colors based on fold change and p-value
proExp = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_ars-dmso.rds', sep = '')))
proExp$logPValue = -log10(proExp$sca.P.Value)
proExp$logPValueScaled = ifelse(proExp$logPValue > 30, 30, proExp$logPValue)
proExp$pColors = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 1, brewer.pal(4,'OrRd')[4], 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -1, brewer.pal(4,'GnBu')[4], brewer.pal(3,'Greys')[2]))
  
#assign text labels to specific genes of interest
#targetGenes = c('DLG2','PRKCB','LOX')
#proExp$pText = ifelse(proExp$symbol %in% targetGenes, proExp$symbol, '')
proExp$pText = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 2, proExp$symbol, 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -3, proExp$symbol, ''))


#create the plot and save it
ggplot(proExp, aes(logFC, logPValueScaled)) +
  geom_point(size = 0.5, color = proExp$pColors) +
  labs(x = 'log2(arsenite-dmso)', y = '-log10(Adjusted p-value)', title = 'Differential abundance') +
  geom_text_repel(label = proExp$pText, nudge_x = 0, nudge_y = -2, max.overlaps = 55000, size = 2, segment.size = 0.25) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-20,20,2)) +
  scale_y_continuous(limits = c(0,30), breaks = seq(0,50,5)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/scatter_deqms_arsenite-dmso.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)


##read in the data for arsenite and plot
##assign colors based on fold change and p-value
proExp = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_combo-dmso.rds', sep = '')))
proExp$logPValue = -log10(proExp$sca.P.Value)
proExp$logPValueScaled = ifelse(proExp$logPValue > 30, 30, proExp$logPValue)
proExp$pColors = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 1, brewer.pal(4,'OrRd')[4], 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -1, brewer.pal(4,'GnBu')[4], brewer.pal(3,'Greys')[2]))
  
#assign text labels to specific genes of interest
#targetGenes = c('DLG2','PRKCB','LOX')
#proExp$pText = ifelse(proExp$symbol %in% targetGenes, proExp$symbol, '')
proExp$pText = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 2, proExp$symbol, 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -5, proExp$symbol, ''))


#create the plot and save it
ggplot(proExp, aes(logFC, logPValueScaled)) +
  geom_point(size = 0.5, color = proExp$pColors) +
  labs(x = 'log2(combo-dmso)', y = '-log10(Adjusted p-value)', title = 'Differential abundance') +
  geom_text_repel(label = proExp$pText, nudge_x = 0, nudge_y = -2, max.overlaps = 55000, size = 2, segment.size = 0.25) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-20,20,2)) +
  scale_y_continuous(limits = c(0,30), breaks = seq(0,50,5)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/scatter_deqms_combo-dmso.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)

##read in the data for arsenite and plot
##assign colors based on fold change and p-value
proExp = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_ms275-dmso.rds', sep = '')))
proExp$logPValue = -log10(proExp$sca.P.Value)
proExp$logPValueScaled = ifelse(proExp$logPValue > 30, 30, proExp$logPValue)
proExp$pColors = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 1, brewer.pal(4,'OrRd')[4], 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -1, brewer.pal(4,'GnBu')[4], brewer.pal(3,'Greys')[2]))
  
#assign text labels to specific genes of interest
#targetGenes = c('DLG2','PRKCB','LOX')
#proExp$pText = ifelse(proExp$symbol %in% targetGenes, proExp$symbol, '')
proExp$pText = ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC >= 2, proExp$symbol, 
                        ifelse(proExp$sca.P.Value <= 0.001 & proExp$logFC <= -2, proExp$symbol, ''))


#create the plot and save it
ggplot(proExp, aes(logFC, logPValueScaled)) +
  geom_point(size = 0.5, color = proExp$pColors) +
  labs(x = 'log2(ms275-dmso)', y = '-log10(Adjusted p-value)', title = 'Differential abundance') +
  geom_text_repel(label = proExp$pText, nudge_x = 0, nudge_y = -2, max.overlaps = 55000, size = 2, segment.size = 0.25) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-20,20,2)) +
  scale_y_continuous(limits = c(0,30), breaks = seq(0,50,5)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.001), linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/scatter_deqms_ms275-dmso.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)

##last make a plot to compare the the arsenite and combo conditions
cond1 = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_ars-dmso.rds', sep = ''))) %>%
  dplyr::select(accession, symbol, logFC, sca.P.Value) %>%
  dplyr::rename(cond1_logFC = logFC, cond1_sca.P.Value = sca.P.Value)
cond2 = as.data.frame(readRDS(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/dataset_deqms_combo-dmso.rds', sep = ''))) %>%
  dplyr::select(accession, symbol, logFC, sca.P.Value) %>%
  dplyr::rename(cond2_logFC = logFC, cond2_sca.P.Value = sca.P.Value)
proExp = cond1 %>%
  full_join(cond2)
proExp$pColors = ifelse(proExp$cond1_logFC >= 1 & proExp$cond2_logFC >= 1, brewer.pal(4,'OrRd')[4], 
                        ifelse(proExp$cond1_logFC <= -1 & proExp$cond2_logFC <= -1, brewer.pal(4,'GnBu')[4], brewer.pal(3,'Greys')[2]))
  
#assign text labels to specific genes of interest
proExp$pText = ifelse(proExp$cond1_logFC >= 2 & proExp$cond2_logFC >= 2, proExp$symbol, 
                        ifelse(proExp$cond1_logFC <= -3 & proExp$cond2_logFC <= -3, proExp$symbol,
                               ifelse(proExp$cond1_logFC <= -2 & proExp$cond2_logFC >= -1, proExp$symbol, 
                                      ifelse(proExp$cond2_logFC <= -3 & proExp$cond1_logFC >= -1, proExp$symbol, ''))))


#create the plot and save it
ggplot(proExp, aes(cond1_logFC, cond2_logFC)) +
  geom_point(size = 0.5, color = proExp$pColors) +
  labs(x = 'log2(arsenite-dmso)', y = 'log2(combo-dmso)', title = 'Differential abundance') +
  geom_text_repel(label = proExp$pText, nudge_x = 0, nudge_y = -2, max.overlaps = 55000, size = 2, segment.size = 0.25) +
  scale_x_continuous(limits = c(-6,6), breaks = seq(-20,20,2)) +
  scale_y_continuous(limits = c(-6,6), breaks = seq(-20,20,2)) +
  geom_vline(xintercept = c(-1,1), linetype = 'dashed') +
  geom_hline(yintercept = c(-1,1), linetype = 'dashed') +
  theme_classic()
ggsave(paste(baseRepository, '/proteomics20230223_a673p1Ybx1HibitTotalProteome/scatter_deqms_arseniteVsComboVsDmso.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```





