---
title: "sorensenTimecourseRna"
author: "Christopher Hughes"
format: html
---

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| label: load-packages and set directories

##########################################################################################
#packages
library('tidyverse')
library('RColorBrewer')
library('ggplot2')
library('tximport')
library('GenomicFeatures')
library('DESeq2')
library('org.Hs.eg.db')
library('ggrepel')

##########################################################################################
#directories
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
```

## Data processing

Read in the sample annotation details and make the txdb that we can use later on for annotation gene identifiers.

```{r}
#| label: read input data
#| warning: false

##########################################################################################
#read a table containing the sample information for our RNAseq runs
samples = read_tsv(paste(baseRepository, '/sequencing20210421_a673TimecourseRnaSeqOutput/sampleInfo.txt', sep = ''))
samples

#use the sample info to build a file list
files = file.path(baseRepository, 'sequencing20210421_a673TimecourseRnaSeqOutput', 'quants', paste(samples$barcode, '_', samples$batch, sep = ''), 'quant.sf')
all(file.exists(files))
```

For Salmon analysis, I am generally following the documentation found [here](http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html) and [here](http://bioconductor.org/packages/release/bioc/vignettes/tximport/inst/doc/tximport.html). When we import our data, we want a table that allows us to link gene and transcript identifiers. For this we use the GTF associated with our database files that we used during the alignment process.

```{r}
#| label: build txdb
#| warning: false

##########################################################################################
#build the txdb from the gtf file
myTxdb = makeTxDbFromGFF('D:/databases/projectEwsDlg2/baseGenomeFiles/genome.gtf')
k = keys(myTxdb, keytype = 'TXNAME')
tx2gene = AnnotationDbi::select(myTxdb, k, 'GENEID', 'TXNAME')
head(tx2gene)
```

Read in the Salmon data.

```{r}
#| label: read salmon data
#| warning: false

##########################################################################################
#read the salmon data
txi = tximport(files, 
               type = 'salmon', 
               tx2gene = tx2gene)
names(txi)
head(txi$counts)
```

```{r}
#| label: perform deseq analysis
#| warning: false

##########################################################################################
#perform the deseq analysis
ddsTxi = DESeqDataSetFromTximport(txi,
                                  colData = samples,
                                  design = ~ sampleName)
dds = DESeq(ddsTxi)
keep = rowSums(counts(dds)) >= 10
dds = dds[keep,]
```

Extract the DESeq data and plot for the different comparisons of interest.

```{r}
#| label: extract and plot comparisons
#| warning: false

##########################################################################################
#parse the deseq data - change the first two values to get different comparisons
daysToCompare = c('day0','day9','day10','day11','day14','day17','day22')
for (i in 1:length(daysToCompare)){
  datasetFirst = daysToCompare[i]
  datasetSecond = 'day7'
  res = results(dds, contrast = c('sampleName', datasetFirst, datasetSecond))
  ens.str = substr(rownames(res), 1, 15)
  res$symbol = mapIds(org.Hs.eg.db,
                      keys=ens.str,
                      column="SYMBOL",
                      keytype="ENSEMBL",
                      multiVals="first")
  resOrdered = res[order(res$pvalue),]
  
  #save the data
  outputData = as.data.frame(resOrdered) %>%
    dplyr::filter(!is.na(symbol)) %>%
    dplyr::select(symbol, log2FoldChange) %>%
    dplyr::rename(sorensenTimecourseRna = log2FoldChange) %>%
    dplyr::group_by(symbol) %>%
    dplyr::summarise(sorensenTimecourseRna = mean(sorensenTimecourseRna, na.rm = TRUE))
  #
  saveRDS(outputData, 
          paste(baseRepository, '/database20220929_ewsExpressionDatabase/sorensenTimecourseRna/dataset_sorensenTimecourseRna_', datasetFirst, '-', datasetSecond, '.rds', sep = ''))
  write.csv(outputData, 
            file = paste(baseRepository, '/database20220929_ewsExpressionDatabase/sorensenTimecourseRna/dataset_sorensenTimecourseRna_', datasetFirst, '-', datasetSecond, '.csv', sep = ''))
}
```

We are done here for now. Wrap up below.

### Session info

```{r}
##########################################################################################
sessionInfo()
```
