---
title: "Polysome curve plotting"
output:
  html_notebook:
      code_folding: none
---

This document details plotting of curves from polysome fractionation. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning = FALSE}
###########################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
library('zoo')
```

<div style="margin-bottom:50px;"></div>

Set the base folders where we will do the work.

```{r}
###########################################################
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToOthers'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToOthers'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read the curve file generated by the fractionator.

```{r, message = FALSE, warning = FALSE}
###########################################################
polyData = read_csv(paste(baseRepository, '/polysomes20210413_mikeOsteoCr131bTreatment/ch20210413_mikeOsteoCr131bTreatment_sample6.csv', sep = ''), skip = 45) %>%
  dplyr::rename(position = `Position(mm)`, fractionNumber = `Fraction Number`, absorbance = Absorbance) %>%
  dplyr::select(position, absorbance, fractionNumber) %>%
  dplyr::filter(position > 10 & position < 55) %>%
  dplyr::mutate(scaledAbsorbance = (absorbance / sum(absorbance)*100))

##
ggplot(polyData, aes(position, scaledAbsorbance)) +
  geom_line(size = 0.5) +
  scale_y_continuous(limits = c(0,0.4), breaks = seq(0,2,0.1)) +
  labs(x = 'Position (mm)', y = 'Normalized absorbance', title = 'Polysome profile') +
  theme_classic()
ggsave(paste(baseRepository, '/polysomes20210413_mikeOsteoCr131bTreatment/line_polysomeDataNormalized_sample6.pdf', sep = ''),
       height = 4, width = 3, useDingbats = FALSE)
```




## Session Info

```{r}
###########################################################
sessionInfo()
```








