---
title: "DLG2 in GTEx"
author: "Christopher Hughes"
format: html
---

## Details

This document describes processing of RNA-seq data from the GTEx project to investigate patterns of DLG2 expression across different tissue types.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| message: false
#| warning: false
##########################################################################################
library('tidyverse')
library('vroom')
library('RColorBrewer')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToEwsCrisprManuscript'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToEwsCrisprManuscript'
```

## Data processing

Read the sample annotation details.

```{r}
##########################################################################################
sampleInfo = read_tsv(paste(generalDatasets, '/gtexVersion8/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt', sep = ''), show_col_types = FALSE)
```

Read in the rna data and reshape it. This is a huge dataset, so you just have to take your gene of interest individually.

```{r}
##########################################################################################
rna = vroom(paste(generalDatasets, '/gtexVersion8/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_tpm.gct', sep = ''), skip = 2, delim = '\t', show_col_types = FALSE) %>%
  filter(grepl('DLG2$', Description)) %>%
  pivot_longer(cols = `GTEX-1117F-0226-SM-5GZZ7`:`GTEX-ZZPU-2726-SM-5NQ8O`, names_to = 'SAMPID', values_to = 'readCount') %>%
  filter(readCount >= 1) %>%
  mutate(log2Tpm = log2(readCount)) %>%
  left_join(sampleInfo)
```

Make a plot for DLG2 across the SMTS column.

```{r}
##########################################################################################
ggplot(rna, aes(x = reorder(SMTS, log2Tpm, median, na.rm = TRUE), log2Tpm)) +
  geom_boxplot(size = 0.5, outlier.shape = NA) +
  labs(x = 'Tissue Specification in GTEx', y = 'log2(Transcripts per million)', title = 'DLG2 RNA expression in GTEx') +
  scale_y_continuous(limits = c(0,8), breaks = seq(0,8,2)) +
  geom_hline(yintercept = 2, linetype = 'dashed') +
  theme_classic() +
  #theme(axis.text.x = element_blank())
  theme(axis.text.x = element_text(angle = 45, size = 6, hjust = 1))
ggsave(paste(baseRepository, '/investigatingDlg2/gtexNormalRnaSeq/plots/boxplot_dlg2RnaExpressionSmtsGtex.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```

Consolidate all low expression cases into a single box for easier visualization.

```{r}
##########################################################################################
tissuesOfInterest = c('Brain', 'Pituitary', 'Skin', 'Adrenal Gland', 'Thyroid')
rnaSubset = rna %>%
  dplyr::mutate(tissueClass = factor(ifelse(SMTS %in% tissuesOfInterest, SMTS, 'Other'), levels = c('Brain','Pituitary','Skin','Adrenal Gland','Thyroid','Other')))

##
ggplot(rnaSubset, aes(x = reorder(tissueClass, -log2Tpm, median, na.rm = TRUE), log2Tpm, color = tissueClass)) +
  geom_violin(size = 0.25, width = 0.75) +
  geom_boxplot(size = 0.25, outlier.size = 0.25, width = 0.1) +
  labs(x = 'Tissue Specification in GTEx', y = 'log2(TPM + 1)', title = 'DLG2 in GTEx') +
  scale_color_manual(values = c(rep(brewer.pal(4,'OrRd')[4],5), brewer.pal(3,'Greys')[3])) +
  scale_y_continuous(limits = c(0,8), breaks = seq(0,8,2)) +
  geom_hline(yintercept = 2, linetype = 'dashed') +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/investigatingDlg2/gtexNormalRnaSeq/plots/boxplot_dlg2RnaExpressionSmtsGtexSubset.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```

