---
title: "soft agar analysis"
author: "Christopher Hughes"
format: html
---

## Details

This document describes processing soft agar colony counting data for DLG2 KO cells.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| message: false
#| warning: false
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
```

## Data processing

Read in the soft agar data and shape into a usable format for plotting. One of the wells for A673p1 was not usable, so we only have 2 replicates there.

```{r}
##########################################################################################
##get the names of all of the colony counting files
sampleFiles = list.files(paste(baseRepository,'/softAgar20230314_a673p1Dlg2Ko8-11-CW', sep = ''), pattern = '*.txt', full.names = TRUE)

##read in the files
sampleData = tibble()
for (i in 1:length(sampleFiles)){
  tempData = read_tsv(sampleFiles[i], show_col_types = FALSE, skip = 2) %>%
    dplyr::select(`Colony Number`, Area, Volume) %>%
    dplyr::rename(colonyNumber = `Colony Number`) %>%
    dplyr::mutate(sampleName = sub('.*CW\\/(.*)_.*','\\1',sampleFiles[i]),
                  replicate = sub('.*CW\\/.*_(.*)\\.txt', '\\1',sampleFiles[i])) 
  #
  sampleData = rbind(sampleData, tempData)
}
```

Make a plot of the data.

```{r}
##########################################################################################
##colony counts
sampleCounts = as.data.frame(table(sampleData$sampleName,sampleData$replicate)) %>%
  dplyr::rename(cellLine = 'Var1', replicate = 'Var2', colonyCount = 'Freq') %>%
  dplyr::filter(colonyCount > 1) %>%
  dplyr::mutate(cellLine = factor(cellLine, levels = c('a673p1', 'ko8', 'ko11', 'pCW-Dlg2s'))) %>%
  dplyr::group_by(cellLine) %>%
  dplyr::summarise(ccMean = median(colonyCount), ccSd = sd(colonyCount))




ggplot(sampleCounts, aes(x = cellLine, y = ccMean, fill = cellLine)) +
  geom_bar(stat = 'identity', width = 0.5) +
  geom_errorbar(aes(ymin=ccMean-ccSd, ymax=ccMean+ccSd), width = 0.2) +
  scale_fill_manual(values = c(brewer.pal(3,'Greys')[2], rep(brewer.pal(4,'GnBu')[4],2), brewer.pal(4,'OrRd')[4])) +
  labs(x = 'Cell line', y = 'Median colony count', title = 'Colony formation') +
  scale_y_continuous(limits = c(0,200), breaks = seq(0,200,50)) +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/softAgar20230314_a673p1Dlg2Ko8-11-CW/barplot_colonyCounts.pdf', sep = ''),
         height = 2, width = 2, useDingbats = FALSE)
```


