---
title: "Analysis of SP3 PRD DNA and RNA binding"
output:
  html_notebook:
      code_folding: none
---

This document is simply for making a plot of the qPCR results from the DNA and RNA yields after SP3 processing. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, echo = FALSE}
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToSp3Prd'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToSp3Prd'
```

<div style="margin-bottom:50px;"></div>

## Data processing

I have the data recorded in my lab book, so we will just input them manually here.

```{r}
##########################################################################################
qpcr = read_tsv(paste(baseRepository, '/purification20200624_Sp3DnaRnaPurificationYields/ch_20200627_QiaSp3PrdPurification_DnaRnaAnalysis_Results.txt', sep = ''))
```

<div style="margin-bottom:50px;"></div>

Parse out only the data we need.

```{r}
##########################################################################################
qpcrSubset = qpcr %>%
  dplyr::select(`Sample Name`, CT) %>%
  mutate(method = sub('.*na(.*)[0-9]+.*', '\\1', `Sample Name`)) %>%
  mutate(molecule = substr(`Sample Name`, 1, 3)) %>%
  mutate(replicate = substr(`Sample Name`, nchar(`Sample Name`), nchar(`Sample Name`))) %>%
  dplyr::select(method, molecule, replicate, CT)
```

<div style="margin-bottom:50px;"></div>

Plot the dna data.

```{r}
##########################################################################################
dna = filter(qpcrSubset, molecule == 'Dna') %>%
  group_by(method, replicate) %>%
  summarise(meanSignal = mean(CT))
##
ggplot(dna, aes(method, meanSignal, color = method)) +
  #geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', width = 0.25) + 
  geom_point(size = 4, position = position_jitterdodge(jitter.width = 0.25)) +
  labs(x = 'Method', y = 'qPCR Measurement (CT)', title = 'DNA signal') +
  scale_y_continuous(limits = c(0, 30), breaks = seq(0,100,10)) +
  scale_color_manual(values = c(brewer.pal(8,'Greys')[4], brewer.pal(8,'RdBu')[1])) +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/purification20200624_Sp3DnaRnaPurificationYields/scatter_qiaSp3DnaQpcr.pdf', sep = ''),
       height = 4,
       width = 2,
       useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Calculate a p-value for the plot.

```{r}
##########################################################################################
t.test(meanSignal ~ method, data = dna)
```

<div style="margin-bottom:50px;"></div>

Now for rna.

```{r}
##########################################################################################
rna = filter(qpcrSubset, molecule == 'Rna') %>%
  group_by(method, replicate) %>%
  summarise(meanSignal = mean(CT))
##
ggplot(rna, aes(method, meanSignal, color = method)) +
  #geom_bar(position = 'dodge', stat = 'summary', fun = 'mean', width = 0.25) + 
  geom_point(size = 4, position = position_jitterdodge(jitter.width = 0.25)) +
  labs(x = 'Method', y = 'qPCR Measurement (CT)', title = 'RNA signal') +
  scale_y_continuous(limits = c(0, 20), breaks = seq(0,100,10)) +
  scale_color_manual(values = c(brewer.pal(8,'Greys')[4], brewer.pal(8,'RdBu')[1])) +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/purification20200624_Sp3DnaRnaPurificationYields/scatter_qiaSp3RnaQpcr.pdf', sep = ''),
       height = 4,
       width = 2,
       useDingbats = FALSE)
```
<div style="margin-bottom:50px;"></div>

Calculate a p-value for the plot.

```{r}
##########################################################################################
t.test(meanSignal ~ method, data = rna)
```

## Session info

```{r}
sessionInfo()
```
