---
title: "ybx1 hibit screening"
  html_notebook:
      code_folding: none
---

This document details the analysis of some plate reader data to investigate HiBit insertion in A673 cells. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Make a plot for the yb1 data. There are only two samples and the output from the plate reader is a bit odd, so I will just input the data manually.

```{r}
##########################################################################################
##data input
hibit = tibble(sample = c('blank','ybx1Neg','ybx1Pos'),
               lum = c(0,411358,210366))

##make the plot
ggplot(hibit, aes(sample, lum)) +
  geom_bar(stat = 'identity', width = 0.75) +
  labs(x = 'Sample', y = 'Luminescence', title = 'HiBit in EwS') +
  theme_classic()
ggsave(paste(baseRepository, '/plateReader20220530_hibitDlg2Ybx1Parentals/barplot_ybx1HiBitSignal.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```

<div style="margin-bottom:50px;"></div>

Make a plot for the dlg2 data. 

```{r}
##########################################################################################
##data input
hibit = tibble(sample = c('blank','dlg2Neg','dlg2Pos'),
               lum = c(0,52,914))

##make the plot
ggplot(hibit, aes(sample, lum)) +
  geom_bar(stat = 'identity', width = 0.75) +
  labs(x = 'Sample', y = 'Luminescence', title = 'HiBit in EwS') +
  theme_classic()
ggsave(paste(baseRepository, '/plateReader20220530_hibitDlg2Ybx1Parentals/barplot_dlg2HiBitSignal.pdf', sep = ''),
       height = 2, width = 2, useDingbats = FALSE)
```

