---
title: "HiBit RIP qPCR"
author: "Christopher Hughes"
format: html
---

This document details the analysis of qPCR data for cells expressing the YBX1 HiBit fusion.

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r}
#| message: false
#| warning: false
##########################################################################################
library('tidyverse')
library('RColorBrewer')
library('ggplot2')
```

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

## Data processing

All I really want to do here is make a plot for the qPCR data. I will input the values manually.

```{r}
##########################################################################################
##make the data table
qpcr = tibble('target' = rep(c('YBX1','G3BP1','HIF1A','NRF2'), each = 3),
              'foldChange' = c(9.84, 9.84, 10.10,
                               8.75, 9.03, 9.19,
                               8.39, 8.49, 8.61,
                               9.43, 9.56, 9.67),
              'replicate' = rep(seq(1,3,1),4)) %>%
  dplyr::mutate(target = factor(target, levels = c('YBX1','G3BP1','HIF1A','NRF2')))

##make the plot
ggplot(qpcr) +
  geom_bar(aes(target, foldChange, fill = target), position = 'dodge', stat = 'summary', fun = 'mean') +
  geom_point(aes(target, foldChange), position = position_jitter(width = 0.5)) +
  labs(x = 'Target', y = 'log2(Fold change relative to parent)', title = 'YB-1 FLAG RIP') +
  theme_classic() +
  theme(legend.position = 'none')
ggsave(paste(baseRepository, '/qpcr20220805_a673HibitRnaIp/barplot_ybx1RipHibitQpcr.pdf', sep = ''),
         height = 2, width = 2, useDingbats = FALSE)

```