---
title: "Analysis of DLG2 IPMS Samples"
output:
  html_notebook:
      code_folding: none
---

This document details the analysis of IPMS samples using DLG2 as a bait (268 antibody). This was done across multiple Ewing sarcoma cell lines, using HOS osteosarcoma cells (isogenic clone A) as a control population (negative for DLG2).

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message=FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
library('ggplot2')
library('RColorBrewer')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
#generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
#baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToDlg2'
baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToDlg2'
```

<div style="margin-bottom:50px;"></div>

## Data processing

Read in the psm data and process into a peptide set.

```{r}
##########################################################################################
psm = read_tsv(paste(baseRepository, '/proteomics20210726_dlg2IpmsEwsCells/combined_peptide.tsv', sep = '')) %>%
  dplyr::select(`Protein ID`, Gene, Sequence, contains('Intensity')) %>%
  dplyr::rename(accession = `Protein ID`,
                gene = Gene,
                sequence = Sequence,
                a673_1 = `A673EF1_1 Intensity`,
                a673_2 = `A673EF1_2 Intensity`,
                a673_3 = `A673EF1_3 Intensity`,
                chla10a_1 = `CHLA10a_1 Intensity`,
                chla10a_2 = `CHLA10a_2 Intensity`,
                chla10a_3 = `CHLA10a_3 Intensity`,
                hosa_1 = `HOSa_1 Intensity`,
                hosa_2 = `HOSa_2 Intensity`,
                hosa_3 = `HOSa_3 Intensity`,
                sknmca_1 = `SKNMCa_1 Intensity`,
                sknmca_2 = `SKNMCa_2 Intensity`,
                sknmca_3 = `SKNMCa_3 Intensity`) %>%
  dplyr::select(accession, gene, sequence, hosa_1:hosa_3, chla10a_1:chla10a_3, sknmca_1:sknmca_3, a673_1:a673_3)
```

<div style="margin-bottom:50px;"></div>

Now that the data are in peptide format, I want to do the individual comparisons. I think it is better to do this cell line by cell line because they may have very different abundances of the bait, or other proteins that come down with it, so it is not good to treat them as replicates in my opinion. For this, I will write a function that creates output files for SAINT that can be used for comparisons.

```{r}
##########################################################################################
controlCellLine='hosa'
targetCellLine='chla10a'
baitProtein='DLG2'

saintProcess = function(targetCellLine, controlCellLine, baitProtein, ...){
  controlNames = c(paste('IGG_',seq(1,3,1), sep=''))
  targetNames = c(paste(baitProtein,'_',seq(1,3,1), sep=''))
  pepData = psm %>%
    dplyr::select(accession, gene, sequence, contains(controlCellLine), contains(targetCellLine)) %>%
    dplyr::rename_at(vars(contains(controlCellLine)), ~controlNames) %>%
    dplyr::rename_at(vars(contains(targetCellLine)), ~targetNames) %>%
    #dplyr::mutate_at(vars(contains('control')), ~na_if(., . < 0)) %>%
    #dplyr::mutate_at(vars(contains('target')), ~na_if(., . < 0)) %>%
    dplyr::filter(if_all(contains(baitProtein), ~ . > 1)) %>%
    dplyr::mutate(Protein = paste(accession,'|',gene, sep=''), Peptide = sequence) %>%
    dplyr::select(Protein, Peptide, controlNames[1]:targetNames[3])
  ##
  write.table(pepData, paste(baseRepository, 
                             '/proteomics20210726_dlg2IpmsEwsCells/saintInput_', 
                             targetCellLine, 
                             baitProtein, 
                             '.tsv', 
                             sep = ''),
              col.names = TRUE, row.names = FALSE, sep = '\t', quote = FALSE)
}


##
targetCellLines = c('chla10a','sknmca','a673')
lapply(targetCellLines, saintProcess, controlCellLine='hosa', baitProtein='DLG2')
```

<div style="margin-bottom:50px;"></div>

I ran in SAINTq (version 0.0.4, https://sourceforge.net/projects/saint-apms/files/) with the parameter file as below.

```
### SAINTq parameter file
## use # to mark a line as comment

## normalize control intensities
normalize_control=false

## name of file with intensities
input_filename=saintInput_chla10aDLG2.tsv

## type of intensity
## valid: protein, peptide, fragment
input_level=peptide

## column names
protein_colname=Protein
pep_colname=Peptide


## control bait selection rules
compress_n_ctrl=100

## test bait replicate selection rules
compress_n_rep=100

## peptide selection rules
min_n_pep=2
best_prop_pep=0.5
```

<div style="margin-bottom:50px;"></div>

Read the SAINTq results in for the different cell lines and look for preys that are enriched in multiple cell systems.

```{r}
##########################################################################################
chla = read_tsv(paste(baseRepository, '/proteomics20210726_dlg2IpmsEwsCells/saintOutput_chla10aDLG2.tsv', sep = '')) %>%
  dplyr::mutate(accession = sub('(.*)\\|.*$', '\\1', Prey), gene = sub('.*\\|(.*)$', '\\1', Prey)) %>%
  dplyr::rename(chla10NumPep = `#Pep`, chla10AvgP = AvgP, chla10BFDR = BFDR) %>%
  dplyr::select(accession, gene, chla10NumPep, chla10AvgP, chla10BFDR)

sknmc = read_tsv(paste(baseRepository, '/proteomics20210726_dlg2IpmsEwsCells/saintOutput_sknmcaDLG2.tsv', sep = '')) %>%
  dplyr::mutate(accession = sub('(.*)\\|.*$', '\\1', Prey), gene = sub('.*\\|(.*)$', '\\1', Prey)) %>%
  dplyr::rename(sknmcNumPep = `#Pep`, sknmcAvgP = AvgP, sknmcBFDR = BFDR) %>%
  dplyr::select(accession, gene, sknmcNumPep, sknmcAvgP, sknmcBFDR)

a673 = read_tsv(paste(baseRepository, '/proteomics20210726_dlg2IpmsEwsCells/saintOutput_a673DLG2.tsv', sep = '')) %>%
  dplyr::mutate(accession = sub('(.*)\\|.*$', '\\1', Prey), gene = sub('.*\\|(.*)$', '\\1', Prey)) %>%
  dplyr::rename(a673NumPep = `#Pep`, a673AvgP = AvgP, a673BFDR = BFDR) %>%
  dplyr::select(accession, gene, a673NumPep, a673AvgP, a673BFDR)


##
allCells = chla %>%
  left_join(sknmc) %>%
  left_join(a673) %>%
  dplyr::filter(if_all(contains('NumPep'), ~ . > 1)) %>%
  dplyr::filter(if_all(contains('AvgP'), ~ . > 0.5))
write.table(allCells, paste(baseRepository, 
                             '/proteomics20210726_dlg2IpmsEwsCells/dataset_dlg2AllCellLinesSaint.tsv',
                             sep = ''),
              col.names = TRUE, row.names = FALSE, sep = '\t', quote = FALSE)
```

<div style="margin-bottom:50px;"></div>

## Session info

```{r}
##########################################################################################
sessionInfo()
```