---
title: "haifeng crispr library"
  html_notebook:
      code_folding: none
---

This document details the analysis of crispr library data for Haifeng. 

## Setting up the environment

These are packages you will need for this notebook. For exact versions used, please refer to the session info at the bottom of this notebook.

```{r, message = FALSE, warning=FALSE}
##########################################################################################
library('tidyverse')
```

<div style="margin-bottom:50px;"></div>

I want to set a base directory that we can use as a link to the directory where we will do most of the work. I use two directories here because the Workspace is what is pushed to GitHub and contains scripts and plot files, but the Repository is where more of the big data is stored that does not get pushed.

```{r}
##########################################################################################
generalDatasets = 'C:/Users/chughes/Documents/bccrc/projectsRepository/generalDatasets'
baseWorkspace = 'C:/Users/chughes/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToOthers'
baseRepository = 'C:/Users/chughes/Documents/bccrc/projectsRepository/sorensenLab/relatedToOthers'
#generalDatasets = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/generalDatasets'
#baseWorkspace = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsWorkspace/sorensenLab/relatedToYbx1'
#baseRepository = 'C:/Users/chris/OneDrive/Documents/bccrc/projectsRepository/sorensenLab/relatedToYbx1'
```

<div style="margin-bottom:50px;"></div>

## Data processing

First we need to prepare the library annotation file. I think he has used the library found [here](https://www.addgene.org/pooled-library/sabatini-human-crispr-metabolic-knockout/).

```{r}
##########################################################################################
##read the library files
grna = read_tsv(paste(baseRepository, '/sequencing20220603_haifengCrisprScreen/metabolismsgrnas.txt',sep='')) %>%
  dplyr::rename(ID = 'sgRNA', sgRNA = 'sequence') %>%
  dplyr::mutate(Gene_symbol = sub('sg(.*)_[0-9]+','\\1',ID))

##save the file
write.table(grna, paste(baseRepository, '/sequencing20220603_haifengCrisprScreen/library.csv',sep=''),
            col.names = FALSE, row.names = FALSE, sep = ',', quote = FALSE)
```

<div style="margin-bottom:50px;"></div>

Read in the results data.

```{r}
##########################################################################################
##read the files
a673 = read_tsv(paste(baseRepository, '/sequencing20220603_haifengCrisprScreen/hzMetaboA673_rra.gene_summary.txt',sep='')) %>%
  dplyr::select(id,num,`neg|score`,`neg|p-value`,`neg|fdr`,`neg|rank`,`neg|lfc`) %>%
  dplyr::rename(a673_num = num, a673_score = `neg|score`, a673_pval = `neg|p-value`,
                a673_fdr = `neg|fdr`, a673_rank = `neg|rank`, a673_lfc = `neg|lfc`)

#
b20r = read_tsv(paste(baseRepository, '/sequencing20220603_haifengCrisprScreen/hzMetaboB20r_rra.gene_summary.txt',sep='')) %>%
  dplyr::select(id,num,`neg|score`,`neg|p-value`,`neg|fdr`,`neg|rank`,`neg|lfc`) %>%
  dplyr::rename(b20r_num = num, b20r_score = `neg|score`, b20r_pval = `neg|p-value`,
                b20r_fdr = `neg|fdr`, b20r_rank = `neg|rank`, b20r_lfc = `neg|lfc`)


##combine the files
allData = a673 %>%
  dplyr::full_join(b20r)\

##write the data
write.table(allData, paste(baseRepository, '/sequencing20220603_haifengCrisprScreen/dataset_mageckCrisprHits.tsv', sep=''),
            col.names = TRUE, row.names = FALSE, quote = FALSE, sep = '\t')

```